### 2024-04-09 HPC/paper.tex
Initial Setup for ACM styled HPC paper

%\documentclass[sigconf, language=german]{acmart}
\documentclass[sigconf]{acmart} %TODO: Choose German or English
\usepackage{booktabs}

\begin{document}

\title{MPI+OpenMP} %TODO: Add Title
\author{Dreßler Justus}
\orcid{0009-0005-0657-5854}
\affiliation{
    \institution{Friedrich Schiller University}
    \department{Faculty of Mathematics and Computer Science}
    \streetaddress{Fürstengraben 1, Main University Building (Universitätshauptgebäude)}
    \postcode{07743}
    \city{Jena}
    \country{Germany}
}
\email{justus.dressler@uni-jena.de}

\keywords{MPI, MPI+OpenMP,MPI+Threads}

% TODO: insert CCSXML http://dl.acm.org/ccs.cfm

\begin{abstract}

    %TODO: add abstract
    Lorem Ipsum dolor sit amet, consectetur adipiscing elit.
    Mauris dictum quam eget orci porttitor ultrices.
    Suspendisse purus nisl, dapibus et rhoncus in, eleifend id quam. Integer volutpat dolor urna, non auctor magna sodales non.
    Ut in fermentum erat, lacinia malesuada urna.
    Nullam et dolor ut purus fermentum rutrum.
    Suspendisse non velit sollicitudin, rutrum est at, commodo sem. Pellentesque viverra tincidunt mauris ut egestas.
    Integer vel orci sapien. Sed ut mi at lacus vehicula fermentum.
    Integer euismod vestibulum auctor.
    Nunc vulputate leo eget aliquam bibendum.

\end{abstract}

\maketitle

%TODO: add content (preferibly in a new tex document)

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris dictum quam eget orci porttitor ultrices. Suspendisse purus nisl, dapibus et rhoncus in, eleifend id quam. Integer volutpat dolor urna, non auctor magna sodales non. Ut in fermentum erat, lacinia malesuada urna. Nullam et dolor ut purus fermentum rutrum. Suspendisse non velit sollicitudin, rutrum est at, commodo sem. Pellentesque viverra tincidunt mauris ut egestas. Integer vel orci sapien. Sed ut mi at lacus vehicula fermentum. Integer euismod vestibulum auctor. Nunc vulputate leo eget aliquam bibendum.

Etiam at cursus velit. Integer pellentesque, magna nec aliquet lobortis, sapien elit tempus ante, at eleifend eros risus sed dui. Suspendisse consectetur sed purus fermentum sollicitudin. Nam id sem et augue cursus luctus. Donec mauris neque, bibendum pellentesque fermentum sit amet, malesuada et augue. In ante enim, aliquam ac scelerisque id, blandit non massa. Donec placerat gravida ipsum et mattis. Cras scelerisque arcu in vulputate vehicula.

Donec nunc tortor, dignissim vitae ultricies et, mollis at turpis. Ut bibendum sit amet magna rutrum condimentum. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia curae; Morbi a vulputate velit. Suspendisse tellus est, ullamcorper in erat quis, iaculis aliquam diam. Aliquam pretium lobortis sollicitudin. In ac massa dictum, semper nisl id, condimentum ligula. Cras accumsan magna sed malesuada faucibus. Nulla in orci hendrerit, convallis tortor sed, vehicula nunc. Donec id magna bibendum, sodales purus non, congue neque. Quisque metus leo, lobortis ut metus a, dignissim lobortis felis.

Fusce quis posuere metus, ac consequat diam. Donec sed venenatis nisl, in ultricies massa. Suspendisse feugiat vestibulum aliquam. Cras consequat venenatis feugiat. Morbi fermentum faucibus mauris eu rutrum. Sed eget scelerisque sem, vel volutpat dui. Nulla a laoreet arcu, vitae varius erat. Proin varius fringilla dui eget dignissim.

Vivamus non ipsum nunc. Donec sit amet erat vitae enim vulputate ultricies nec id mauris. Phasellus consectetur sem aliquet malesuada egestas. Aliquam viverra libero eros, ac scelerisque elit vestibulum non. Integer viverra lacus quis nibh vulputate pharetra. Vestibulum viverra nulla et facilisis luctus. Sed mollis ultrices nulla. Vivamus ut ornare diam.

\end{document}

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Console/Api/PopClientApi.java
Remove Duplicated Repos and Generated PNG's

@@ -1,160 +1,0 @@
-package Console.Api;
-
-import javax.mail.*;
-import javax.mail.internet.MimeMultipart;
-
-import org.jsoup.Jsoup;
-
-import java.io.IOException;
-
-import java.util.*;
-
-//https://kodejava.org/how-do-i-receive-mail-using-pop3/
-public class PopClientApi {
-    // username "Minutenreis@gmx.de";
-    // password "TestPasswordGithub";
-    // server "pop.gmx.net";
-    // port 995;
-
-    Scanner sc;
-
-    public PopClientApi(Scanner sc) {
-        this.sc = sc;
-    }
-
-    public void receiveMail() throws Exception {
-        // 1. Setup properties for the mail session.
-        System.out.println("Enter your POP3 Server: ");
-        String server = sc.nextLine();
-        System.out.println("Enter your POP3 Port: ");
-        int port = Integer.parseInt(sc.nextLine());
-        System.out.println("Enter your email address: ");
-        String username = sc.nextLine();
-        //https://stackoverflow.com/questions/12076165/how-to-obscure-scanner-input-text
-        String password = new String(System.console().readPassword("Enter your password: "));
-        Properties props;
-        if (port == 995) {
-            props = configPropertiesSSL(server, port, username);
-        } else if (port == 110) {
-            props = configPropertiesNoSSL(server, port, username);
-        } else {
-            System.out.println("SSL? (y/n)");
-            String ssl = sc.nextLine();
-            props = ssl == "n" ? configPropertiesNoSSL(server, port, username)
-                    : configPropertiesSSL(server, port, username);
-        }
-
-        // 2. Creates a javax.mail.Authenticator object.
-        Authenticator auth = new Authenticator() {
-            @Override
-            protected PasswordAuthentication getPasswordAuthentication() {
-                return new PasswordAuthentication(username, password);
-            }
-        };
-
-        // 3. Creating mail session.
-        Session session = Session.getInstance(props, auth);
-
-        // 4. Get the POP3 store provider and connect to the store.
-        Store store = session.getStore("pop3");
-        store.connect(server, port, username, password);
-
-        // 5. Get folder and open the INBOX folder in the store.
-        Folder inbox = store.getFolder("INBOX");
-        inbox.open(Folder.READ_ONLY);
-
-        // 6. Retrieve the messages from the folder.
-        Message[] messages = inbox.getMessages();
-
-        while (true) {
-            // Input message
[…]

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Console/Api/SmtpClientApi.java

@@ -1,128 +1,0 @@
-package Console.Api;
-
-import java.util.*;
-
-import javax.mail.*;
-import javax.mail.internet.InternetAddress;
-import javax.mail.internet.MimeMessage;
-
-public class SmtpClientApi {
-    // username "Minutenreis@gmx.de";
-    // password "TestPasswordGithub";
-    // server "mail.gmx.net";
-    // port 587;
-    Scanner sc;
-
-    public SmtpClientApi(Scanner sc) {
-        this.sc = sc;
-    }
-
-    //https://www.digitalocean.com/community/tutorials/javamail-example-send-mail-in-java-smtp
-    public void sendMail() throws Exception {
-        // 1. Setup properties for the mail session.
-        System.out.println("Enter your SMTP Server: ");
-        String server = sc.nextLine();
-        System.out.println("Enter your SMTP Port: ");
-        int port = Integer.parseInt(sc.nextLine());
-        System.out.println("Enter your email address: ");
-        String username = sc.nextLine();
-        // https://stackoverflow.com/questions/12076165/how-to-obscure-scanner-input-text
-        String password = new String(System.console().readPassword("Enter your password: "));
-
-        System.out.println("Enter recipients (separated by ','): ");
-        String recipients = sc.nextLine();
-        Address[] recipientsArray = InternetAddress.parse(recipients);
-
-        System.out.println("Enter your subject: ");
-        String subject = sc.nextLine();
-        System.out.println("Enter your message line by line. Enter '.' to finish.");
-        String text = "";
-        String nextline;
-        while(!(nextline = sc.nextLine()).equals(".")) {
-            text += nextline + "\n";
-        }
-
-        Properties props = setProperties(server, port, username);
-
-        // have to test whether unencrypted cares for Authenticator
-        // boolean encrypted = props.contains(props.get("mail.smtp.starttls.enable"));
-
-        Session session;
-
-        // if(encrypted){
-            Authenticator auth = new Authenticator() {
-                @Override
-                protected PasswordAuthentication getPasswordAuthentication() {
-                    return new PasswordAuthentication(username, password);
-                }
-            };
-            session = Session.getInstance(props, auth);
-        // } else {
-            // session = Session.getInstance(props);
-        // }
-        
-        // 6. Create a message.
-        Message message = new MimeMessage(session);
-
-        //8. Set Content
-        message.setFrom(new InternetAddress(username));
-        message.setSubject(subject);
-        message.setText(text);
-        message.setSentDate(new Date());
[…]

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Console/Main.java

@@ -1,42 +1,0 @@
-package Console;
-import java.util.Scanner;
-
-import Console.Api.PopClientApi;
-import Console.Api.SmtpClientApi;
-import Console.Socket.PopClientSocket;
-import Console.Socket.SmtpClientSocket;
-
-public class Main {
-    public static void main(String[] args) throws Exception {
-
-        final Scanner scanner = new Scanner(System.in);
-        System.out.println("Receive API: 1\nReceive Socket: 2 \nSend API: 3\nSend Socket: 4");
-        int option = Integer.parseInt(scanner.nextLine());
-        switch (option) {
-            case 1:
-                // Java Mail
-                PopClientApi popClientM = new PopClientApi(scanner);
-                popClientM.receiveMail();
-                break;
-            case 2:
-                // Socket
-                PopClientSocket popClientS = new PopClientSocket(scanner);
-                popClientS.receiveMail();
-                break;
-            case 3:
-                // Java Mail
-                SmtpClientApi smtpClientM = new SmtpClientApi(scanner);
-                smtpClientM.sendMail();
-                break;
-            case 4:
-                // Socket
-                SmtpClientSocket smtpClientS = new SmtpClientSocket(scanner);
-                smtpClientS.sendMail();
-                break;
-            default:
-                System.out.println("Keine Implementation");
-                break;
-        }
-        scanner.close();
-    }
-}

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Console/Socket/BasicMessage.java

@@ -1,35 +1,0 @@
-package Console.Socket;
-
-public class BasicMessage {
-    private final String body;
-    private final String subject;
-    private final String from;
-    private final String[] to;
-
-    protected BasicMessage(String body, String subject, String from, String[] to) {
-        this.body = body;
-        this.subject = subject;
-        this.from = from;
-        this.to = to;
-    }
-
-    public String getBody() {
-        return body;
-    }
-
-    public String getSubject() {
-        return subject;
-    }
-
-    public String getFrom() {
-        return from;
-    }
-
-    public String[] getTo() {
-        return to;
-    }
-
-    public String getToString() {
-        return String.join(",", to);
-    }
-}

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Console/Socket/PopClientSocket.java

@@ -1,172 +1,0 @@
-package Console.Socket;
-
-import java.io.BufferedReader;
-import java.io.IOException;
-import java.io.InputStreamReader;
-import java.io.PrintWriter;
-import java.net.Socket;
-import java.util.Scanner;
-
-import javax.net.ssl.SSLSocketFactory;
-
-// http://mike-java.blogspot.com/2008/03/simple-pop3-client-in-java-tutorial.html
-public class PopClientSocket {
-    Socket socket;
-    private BufferedReader reader;
-    private PrintWriter writer;
-    private boolean debug = false;
-    Scanner sc;
-
-    String testUsername = "Minutenreis@gmx.de";
-    String testPassword = "TestPasswordGithub";
-    String testServer = "pop.gmx.net";
-    int testPort = 995;
-
-    public PopClientSocket(Scanner sc) {
-        this.sc = sc;
-    }
-
-    public String[] receiveMail() throws IOException {
-        debug = true;
-        connect(sc);
-        login(testUsername, testPassword);
-        int stat = getMessageQuantity();
-        System.out.println("Sie haben " + stat + " Nachrichten");
-        while (true) {
-            // Input message
-            System.out.println("Enter message number: ");
-            int messageNumber = Integer.parseInt(sc.nextLine());
-            retrieve(messageNumber, reader);
-            System.out.println("\n Wollen Sie eine andere Nachricht lesen? (y/n)");
-            String input = sc.nextLine();
-            if (input.equals("n")) {
-                break;
-            }
-        }
-        // List<Message> list = getAllMessages();
-        disconnect();
-        socket.close();
-        return null;
-    }
-
-    // Retrieve Message
-    public void retrieve(int messageNumber, BufferedReader reader) throws IOException {
-        if(sendCommand("RETR " + messageNumber).startsWith("-ERR")) {
-            System.out.println("Error retrieving message");
-            return;
-        }
-        boolean text = false;
-        while (true) {
-            String line = reader.readLine();
-            if (line.equals(".")) {
-                break;
-            }
-            if (line.startsWith("Content-Type: text/plain")){
-                text = true;
-            } else if (line.startsWith("--")){
-                text = false;
-            }
-            if(text){
-                System.out.println(line);
-            //important headers
-            } else if(line.startsWith("From:") || line.startsWith("Subject:") || line.startsWith("Date:") || line.startsWith("To:")){
-                System.out.println(line);
-            }
-        }
-    }
-
-    // connect to the server
[…]

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Console/Socket/SmtpClientSocket.java

@@ -1,151 +1,0 @@
-package Console.Socket;
-
-import java.io.BufferedReader;
-import java.io.DataOutputStream;
-import java.io.IOException;
-import java.io.InputStreamReader;
-import java.net.Socket;
-import java.text.DateFormat;
-import java.text.SimpleDateFormat;
-import java.util.ArrayList;
-import java.util.Base64;
-import java.util.Date;
-import java.util.List;
-import java.util.Scanner;
-
-import javax.net.ssl.SSLSocketFactory;
-
-public class SmtpClientSocket {
-    Socket socket;
-    BufferedReader reader;
-    DataOutputStream writer;
-    boolean debug = true;
-    Scanner sc;
-
-    String testUsername = "minutenreis@gmx.de";
-    String testPassword = "TestPasswordGithub";
-    String testServer = "mail.gmx.net";
-    int testPort = 465;
-    boolean ssl = true;
-
-    public SmtpClientSocket(Scanner sc) {
-        this.sc = sc;
-    }
-
-    public String[] sendMail() throws IOException {
-        sendMessage(createMessage(), false);
-        disconnect();
-        socket.close();
-        return null;
-    }
-
-    void connect(String host, int port) throws IOException {
-        socket = ssl ? SSLSocketFactory.getDefault().createSocket(host, port) : new Socket(host, port);
-        reader = new BufferedReader(new InputStreamReader(socket.getInputStream()));
-        writer = new DataOutputStream(socket.getOutputStream());
-        if (debug) {
-            System.out.println("Connected to the " + host + " on port " + port);
-        }
-        readResponseLine();
-    }
-
-    void disconnect() throws IOException {
-        if (!isConnected()) {
-            throw new IllegalStateException("Not connected");
-        }
-        sendCommand("QUIT");
-        socket.close();
-        reader = null;
-        writer = null;
-        if (debug) {
-            System.out.println("Disconnected");
-        }
-    }
-
-    boolean isConnected() {
-        return socket != null && socket.isConnected();
-    }
-
-    String sendCommand(String command) throws IOException {
-        if (debug) {
-            System.out.println(command);
-        }
-        writer.writeBytes(command + "\r\n");
-        return readResponseLine();
-    }
-
-    protected String readResponseLine() throws IOException {
-        String response = reader.readLine();
-        if (debug) {
-            System.out.println(response);
-        }
-        return response;
-    }
-
[…]

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Gui/BtnBar.java

@@ -1,47 +1,0 @@
-import java.awt.event.ActionEvent;
-import java.awt.event.ActionListener;
-
-import javax.swing.JButton;
-import javax.swing.JPanel;
-import javax.swing.JToolBar;
-
-public class BtnBar extends JPanel {
-    
-    JToolBar toolBar = new JToolBar();
-
-    BtnBar(Runnable writeMail, Runnable openDialog, Runnable reloadMails, Runnable deleteMail){
-        JButton btnNewMail = new JButton("New Mail");
-        btnNewMail.addActionListener(new ActionListener(){
-            public void actionPerformed(ActionEvent e){
-                writeMail.run();
-            }
-        });
-        toolBar.add(btnNewMail);
-
-        JButton btnConfig = new JButton("Edit User");
-        btnConfig.addActionListener(new ActionListener(){
-            public void actionPerformed(ActionEvent e){
-                openDialog.run();
-            }
-        });
-        toolBar.add(btnConfig);
-
-        JButton btnReload = new JButton("Reload");
-        btnReload.addActionListener(new ActionListener(){
-            public void actionPerformed(ActionEvent e){
-                reloadMails.run();
-            }
-        });
-        toolBar.add(btnReload);
-
-        JButton btnDelete = new JButton("Delete");
-        btnDelete.addActionListener(new ActionListener(){
-            public void actionPerformed(ActionEvent e){
-                deleteMail.run();
-            }
-        });
-        toolBar.add(btnDelete);
-
-        this.add(toolBar);
-    }
-}

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Gui/DisposablePanel.java

@@ -1,5 +1,0 @@
-import javax.swing.JPanel;
-
-public class DisposablePanel extends JPanel {
-    public Runnable dispose;
-}

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Gui/GuiApi/GuiBasicMessage.java

@@ -1,25 +1,0 @@
-package GuiApi;
-
-public class GuiBasicMessage {
-    public final String body;
-    public final String subject;
-    public final String from;
-    public final String to;
-    public final String cc;
-
-    public GuiBasicMessage(String body, String subject, String from, String to, String cc) {
-        this.body = body;
-        this.subject = subject;
-        this.from = from;
-        this.to = to;
-        this.cc = cc;
-    }
-
-    public String getToString() {
-        return String.join(",", to);
-    }
-
-    public boolean isValid(){
-        return body != null && subject != null && from != null && to != null;
-    }
-}

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Gui/GuiApi/GuiPopClientApi.java

@@ -1,153 +1,0 @@
-package GuiApi;
-
-import java.io.ByteArrayOutputStream;
-import java.io.File;
-import java.io.FileOutputStream;
-import java.io.IOException;
-import java.io.OutputStream;
-import java.util.Properties;
-import java.util.regex.Matcher;
-import java.util.regex.Pattern;
-
-import javax.mail.Authenticator;
-import javax.mail.BodyPart;
-import javax.mail.Folder;
-import javax.mail.Message;
-import javax.mail.MessagingException;
-import javax.mail.PasswordAuthentication;
-import javax.mail.Session;
-import javax.mail.Store;
-import javax.mail.internet.MimeMultipart;
-
-import org.jsoup.Jsoup;
-
-import GuiApi.User.POP3PROTOCOL;
-
-//https://kodejava.org/how-do-i-receive-mail-using-pop3/
-public class GuiPopClientApi {
-    // username "Minutenreis@gmx.de";
-    // password "TestPasswordGithub";
-    // server "pop.gmx.net";
-    // port 995;
-
-    public Message[] receiveMail(User user) throws Exception {
-        // 1. Setup properties for the mail session.
-        Properties props;
-        if (user.pop3Protocol == POP3PROTOCOL.SSL) {
-            props = configPropertiesSSL(user);
-        } else if (user.pop3Protocol == POP3PROTOCOL.NO_SSL) {
-            props = configPropertiesNoSSL(user);
-        } else {
-            throw new Exception("POP3 Protocol not set");
-        }
-
-        // 2. Creates a javax.mail.Authenticator object.
-        Authenticator auth = new Authenticator() {
-            @Override
-            protected PasswordAuthentication getPasswordAuthentication() {
-                return new PasswordAuthentication(user.username, user.password);
-            }
-        };
-
-        // 3. Creating mail session.
-        Session session = Session.getInstance(props, auth);
-
-        // 4. Get the POP3 store provider and connect to the store.
-        Store store = session.getStore("pop3");
-        store.connect(user.pop3Server, user.pop3Port, user.username, user.password);
-
-        // 5. Get folder and open the INBOX folder in the store.
-        Folder inbox = store.getFolder("INBOX");
-        inbox.open(Folder.READ_ONLY);
-
-        // 6. Retrieve the messages from the folder.
-        // int messageCount = inbox.getMessageCount();
-        Message[] messages = inbox.getMessages();
-
-        // 6.5 Write to disk
[…]

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Gui/GuiApi/GuiSmtpClientApi.java

@@ -1,98 +1,0 @@
-package GuiApi;
-
-import java.util.*;
-
-import javax.mail.*;
-import javax.mail.internet.InternetAddress;
-import javax.mail.internet.MimeMessage;
-
-public class GuiSmtpClientApi {
-    // username "Minutenreis@gmx.de";
-    // password "TestPasswordGithub";
-    // server "mail.gmx.net";
-    // port 587;
-
-    // https://www.digitalocean.com/community/tutorials/javamail-example-send-mail-in-java-smtp
-    public void sendMail(User user, GuiBasicMessage basicMessage) throws Exception {
-        if (!basicMessage.isValid()) {
-            throw new Exception("Message is not valid");
-        }
-
-        // 1. Setup properties for the mail session.
-
-        Properties props = setProperties(user);
-
-        Session session;
-
-        if (user.smtpProtocol != User.SMTPPROTOCOL.NO_SSL) {
-            Authenticator auth = new Authenticator() {
-                @Override
-                protected PasswordAuthentication getPasswordAuthentication() {
-                    return new PasswordAuthentication(user.username, user.password);
-                }
-            };
-            session = Session.getInstance(props, auth);
-        } else {
-            session = Session.getInstance(props);
-        }
-
-        // 6. Create a message.
-        Message message = new MimeMessage(session);
-
-        // 8. Set Content
-        message.setFrom(new InternetAddress(user.username));
-        message.setSubject(basicMessage.subject);
-        message.setText(basicMessage.body);
-        message.setSentDate(new Date());
-        message.setRecipients(Message.RecipientType.TO, InternetAddress.parse(basicMessage.to));
-        if (!basicMessage.cc.equals("")) {
-            message.addRecipients(Message.RecipientType.CC, InternetAddress.parse(basicMessage.cc));
-        }
-        // 9. Send the message.
-        Transport.send(message, message.getAllRecipients());
-    }
-
-    Properties configPropertiesNoSSL(User user) {
-        Properties props = new Properties();
-        props = System.getProperties();
-        props.put("mail.smtp.host", user.smtpServer);
-        props.put("mail.smtp.port", user.smtpPort);
-        props.put("mail.smtp.starttls.enable", "false");
-        return props;
-    }
-
-    Properties configPropertiesTLS(User user) {
-        Properties props = new Properties();
-        props = System.getProperties();
-        props.put("mail.smtp.host", user.smtpServer);
[…]

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Gui/GuiApi/User.java

@@ -1,30 +1,0 @@
-package GuiApi;
-
-import java.io.Serializable;
-
-public class User implements Serializable {
-
-    public enum POP3PROTOCOL {
-        SSL, NO_SSL
-    }
-
-    public enum SMTPPROTOCOL {
-        SSL, TLS, NO_SSL
-    }
-    
-    public String username;
-    public String password;
-
-    public String pop3Server;
-    public Integer pop3Port;
-    public POP3PROTOCOL pop3Protocol;
-
-    public String smtpServer;
-    public Integer smtpPort;
-    public SMTPPROTOCOL smtpProtocol;
-
-    public User() {
-        this.pop3Protocol = POP3PROTOCOL.SSL;
-        this.smtpProtocol = SMTPPROTOCOL.TLS;
-    }
-}

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Gui/MainSwing.java

@@ -1,9 +1,0 @@
-import javax.swing.UIManager;
-
-public class MainSwing {
-    public static void main(String[] args) throws Exception {
-        UIManager.setLookAndFeel("com.formdev.flatlaf.FlatDarkLaf");
-        MainWindow mainWindow = new MainWindow();
-        mainWindow.show();
-    }
-}

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Gui/MainWindow.java

@@ -1,364 +1,0 @@
-import java.awt.Color;
-import java.awt.Dialog;
-import java.awt.Dimension;
-import java.awt.FlowLayout;
-import java.awt.Point;
-import java.io.ByteArrayOutputStream;
-import java.io.File;
-import java.io.FileInputStream;
-import java.io.FileOutputStream;
-import java.io.InputStream;
-import java.io.ObjectInputStream;
-import java.io.OutputStream;
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.Comparator;
-import java.util.Date;
-import java.util.Properties;
-import java.util.spi.CurrencyNameProvider;
-
-import javax.mail.Address;
-import javax.mail.Flags;
-import javax.mail.Message;
-import javax.mail.Session;
-import javax.mail.internet.MimeMessage;
-import javax.mail.internet.MimeUtility;
-import javax.swing.JDialog;
-import javax.swing.JFrame;
-import javax.swing.JScrollBar;
-import javax.swing.JScrollPane;
-import javax.swing.JTable;
-import javax.swing.JTextArea;
-import javax.swing.ListSelectionModel;
-import javax.swing.event.ListSelectionEvent;
-import javax.swing.event.ListSelectionListener;
-import javax.swing.table.DefaultTableModel;
-
-import GuiApi.GuiPopClientApi;
-import GuiApi.User;
-
-public class MainWindow {
-
-    private JFrame window;
-    private User user;
-
-    private JTextArea currentMail = new JTextArea(20, 49);
-    private int currentMailIndex = -1;
-
-    private JTable messageTable = new JTable();
-    String[] columns = { "Read", "Sent", "Subject", "From", "To", "FileName" };
-
-    public MainWindow() throws Exception {
-        loadUser();
-
-        window = new JFrame();
-        window.setTitle("Reismail");
-        window.setDefaultCloseOperation(JFrame.DISPOSE_ON_CLOSE);
-        window.setSize(800, 800);
-        window.setLocationRelativeTo(null);
-        window.setVisible(true);
-        window.setLayout(new FlowLayout());
-
-        // Config Dialog
-        if (user == null) {
-            user = new User();
-            UserConfigDialog mandatoryConfigDialog = new UserConfigDialog(user, false);
-            createModalDialog(window, "User Config", mandatoryConfigDialog).setVisible(true);
-        }
-
-        UserConfigDialog configDialog = new UserConfigDialog(user, true);
-
[…]

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Gui/UserConfigDialog.java

@@ -1,170 +1,0 @@
-import java.awt.Component;
-import java.awt.event.ActionEvent;
-import java.awt.event.ActionListener;
-import java.io.File;
-import java.io.FileOutputStream;
-import java.io.ObjectOutputStream;
-
-import javax.swing.BoxLayout;
-import javax.swing.ButtonGroup;
-import javax.swing.JButton;
-import javax.swing.JLabel;
-import javax.swing.JPanel;
-import javax.swing.JPasswordField;
-import javax.swing.JRadioButton;
-import javax.swing.JTextField;
-
-import java.awt.*;
-
-import GuiApi.User;
-import GuiApi.User.POP3PROTOCOL;
-import GuiApi.User.SMTPPROTOCOL;
-
-public class UserConfigDialog extends DisposablePanel {
-
-    public UserConfigDialog(User user, boolean cancelable) {
-        this.setLayout(new BoxLayout(this, BoxLayout.Y_AXIS));
-
-        JLabel usernameLabel = new JLabel("Username");
-        Component[] usernameLabels = { usernameLabel };
-        addSubPanel(usernameLabels);
-
-        JTextField usernameText = new JTextField(user.username != null? user.username : "Minutenreis@gmx.de", 20);
-        usernameText.setMaximumSize(usernameText.getPreferredSize());
-        Component[] usernameTexts = { usernameText };
-        addSubPanel(usernameTexts);
-
-        JLabel passwordLabel = new JLabel("Password");
-        Component[] passwordLabels = { passwordLabel };
-        addSubPanel(passwordLabels);
-
-        JPasswordField passwordText = new JPasswordField(user.password != null? user.password : "TestPasswordGithub", 20);
-        passwordText.setMaximumSize(passwordText.getPreferredSize());
-        Component[] passwordTexts = { passwordText };
-        addSubPanel(passwordTexts);
-
-        JLabel pop3ServerLabel = new JLabel("Pop3-Server");
-        JLabel smtpServerLabel = new JLabel("Smtp-Server");
-
-        Component[] serverLabels = { pop3ServerLabel, smtpServerLabel };
-        addSubPanel(serverLabels);
-
-        JTextField pop3ServerText = new JTextField(user.pop3Server != null? user.pop3Server : "pop.gmx.net", 20);
-        pop3ServerText.setMaximumSize(pop3ServerText.getPreferredSize());
[…]

### 2024-04-09 Archive/WiSe_2022_2023/FortgeschrittenesProgrammierpraktikum/src/Gui/WriteMailDialog.java

@@ -1,108 +1,0 @@
-import java.awt.Component;
-import java.awt.event.ActionEvent;
-import java.awt.event.ActionListener;
-
-import javax.swing.BoxLayout;
-import javax.swing.JButton;
-import javax.swing.JLabel;
-import javax.swing.JPanel;
-import javax.swing.JTextField;
-import javax.swing.JTextArea;
-
-
-import GuiApi.GuiBasicMessage;
-import GuiApi.GuiSmtpClientApi;
-import GuiApi.User;
-
-public class WriteMailDialog extends DisposablePanel {
-
-    public WriteMailDialog(User user, boolean cancelable){
-        this.setLayout(new BoxLayout(this, BoxLayout.Y_AXIS));
-
-
-        // Labels & Textfields 
-        JLabel recipientLabel = new JLabel("Recipient:");
-        Component[] recipientLabelComponents = { recipientLabel };
-        addSubPanel(recipientLabelComponents);
-
-        JTextField recipientText = new JTextField(20);
-        recipientText.setMaximumSize(recipientText.getPreferredSize());
-        Component[] recipientComponents = { recipientText };
-        addSubPanel(recipientComponents);
-
-        JLabel ccLabel = new JLabel("CC:");
-        Component[] ccLabelComponents = { ccLabel };
-        addSubPanel(ccLabelComponents);
-
-        JTextField ccText = new JTextField(20);
-        recipientText.setMaximumSize(ccText.getPreferredSize());
-        Component[] ccComponents = { ccText };
-        addSubPanel(ccComponents);
-
-        JLabel subjectLabel = new JLabel("Subject:");
-        Component[] subjectLabelComponents = { subjectLabel };
-        addSubPanel(subjectLabelComponents);
-
-        JTextField subjectText = new JTextField(20);
-        subjectText.setMaximumSize(subjectText.getPreferredSize());
-        Component[] subjectComponents = { subjectText };
-        addSubPanel(subjectComponents);
-
-        JLabel messageLabel = new JLabel("Message:");
-        Component[] messageLabelComponents = { messageLabel };
-        addSubPanel(messageLabelComponents);
-
-        JTextArea messageText = new JTextArea(20,20);
-        messageText.setMaximumSize(messageText.getPreferredSize());
-        Component[] messageComponents = { messageText };
-        addSubPanel(messageComponents);
-
-
-        // Buttons
-        JButton btnSend = new JButton("send");
-        btnSend.addActionListener(new ActionListener(){
[…]

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_22_10_24_dressler_kroehl_halank/src/binaryRepresentation.cpp

@@ -1,40 +1,0 @@
-#include <iostream>
-#include <bitset>
-
-int main()
-{
-    unsigned char l_data1 = 1;
-    unsigned char l_data2 = 255;
-    unsigned char l_data3 = l_data2 + 1;
-    unsigned char l_data4 = 0xA1;
-    unsigned char l_data5 = 0b1001011;
-    unsigned char l_data6 = 'H';
-    char l_data7 = -4;
-    unsigned int l_data8 = 1u << 11;
-    unsigned int l_data9 = l_data8 << 21;
-    unsigned int l_data10 = 0xFFFFFFFF >> 5;
-    unsigned int l_data11 = 0b1001 ^ 0b01111;
-    unsigned int l_data12 = ~0b1001;
-    unsigned int l_data13 = 0xF0 & 0b1010101;
-    unsigned int l_data14 = 0b001 | 0b101;
-    unsigned int l_data15 = 7743;
-    int l_data16 = -7743;
-    std::cout << "l_data1 = " << std::bitset< 8 >( l_data1 ) << std::endl;
-    std::cout << "l_data2 = " << std::bitset< 8 >( l_data2 ) << std::endl;
-    std::cout << "l_data3 = " << std::bitset< 8 >( l_data3 ) << std::endl;
-    std::cout << "l_data4 = " << std::bitset< 8 >( l_data4 ) << std::endl;
-    std::cout << "l_data5 = " << std::bitset< 8 >( l_data5 ) << std::endl;
-    std::cout << "l_data6 = " << std::bitset< 8 >( l_data6 ) << std::endl;
-    std::cout << "l_data7 = " << std::bitset< 8 >( l_data7 ) << std::endl;
-    std::cout << "l_data8 = " << std::bitset< 32 >( l_data8 ) << std::endl;
-    std::cout << "l_data9 = " << std::bitset< 32 >( l_data9 ) << std::endl;
-    std::cout << "l_data10 = " << std::bitset< 32 >( l_data10 ) << std::endl;
-    std::cout << "l_data11 = " << std::bitset< 32 >( l_data11 ) << std::endl;
-    std::cout << "l_data12 = " << std::bitset< 32 >( l_data12 ) << std::endl;
-    std::cout << "l_data13 = " << std::bitset< 32 >( l_data13 ) << std::endl;
-    std::cout << "l_data14 = " << std::bitset< 32 >( l_data14 ) << std::endl;
[…]

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_22_12_12_dressler_kroehl_halank/src/add_c.c

@@ -1,16 +1,0 @@
-#include <stdint.h>
-
-void add_c(uint64_t i_n_values,
-           uint64_t const *i_a,
-           uint64_t const *i_b,
-           uint64_t *o_c)
-{
-  for (uint64_t l_va = 0; l_va < i_n_values; l_va++)
-  {
-    uint64_t l_tmp_a = i_a[l_va];
-    uint64_t l_tmp_b = i_b[l_va];
-    uint64_t l_tmp_c = l_tmp_a + l_tmp_b;
-
-    o_c[l_va] = l_tmp_c;
-  }
-}

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_22_12_12_dressler_kroehl_halank/src/add_driver.cpp

@@ -1,68 +1,0 @@
-#include <cstdint>
-#include <cstdlib>
-#include <iostream>
-
-extern "C" {
-  void add_c( uint64_t         i_n_values,
-              uint64_t const * i_a,
-              uint64_t const * i_b,
-              uint64_t       * o_c );
-  void add_asm( uint64_t         i_n_values,
-                uint64_t const * i_a,
-                uint64_t const * i_b,
-                uint64_t       * o_c );
-}
-
-int main() {
-  uint64_t l_n_values = 10;
-
-  // init pointers
-  uint64_t * l_a = nullptr;
-  uint64_t * l_b = nullptr;
-  uint64_t * l_c_0 = nullptr;
-  uint64_t * l_c_1 = nullptr;
-
-  // allocate memory
-  l_a   = (uint64_t *) new uint64_t[ l_n_values ];
-  l_b   = (uint64_t *) new uint64_t[ l_n_values ];
-  l_c_0 = (uint64_t *) new uint64_t[ l_n_values ];
-  l_c_1 = (uint64_t *) new uint64_t[ l_n_values ];
-
-  // init arrays
-  for( std::size_t l_va = 0; l_va < l_n_values; l_va++ ) {
-    l_a[l_va] = l_va;
-    l_b[l_va] = l_va*2;
-    l_c_0[l_va] = 0;
-    l_c_1[l_va] = 0;
-  }
-
-  // add_c
-  std::cout << "### calling add_c ###" << std::endl;
-  add_c( l_n_values,
-         l_a,
-         l_b,
-         l_c_0 );
-
-  for( std::size_t l_va = 0; l_va < l_n_values; l_va++ ) {
-    std::cout << l_a[l_va] << " / " << l_b[l_va] << " / " << l_c_0[l_va] << std::endl;
-  }
-
-  // add_asm
-  std::cout << "### calling add_asm ###" << std::endl;
-  add_asm( l_n_values,
-           l_a,
-           l_b,
-           l_c_1 );
-
-  for( std::size_t l_va = 0; l_va < l_n_values; l_va++ ) {
-    std::cout << l_a[l_va] << " / " << l_b[l_va] << " / " << l_c_1[l_va] << std::endl;
-  }
-
-  // free memory
-  delete [] l_a;
-  delete [] l_b;
-  delete [] l_c_0;
-  delete [] l_c_1;
-
-  return EXIT_SUCCESS;
-}

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_22_12_12_dressler_kroehl_halank/src/copy_c.c

@@ -1,20 +1,0 @@
-#include <stdint.h>
-
-void copy_c( uint64_t const * i_a,
-             uint64_t       * i_b ) {
-  uint64_t l_tmp_0 = i_a[0];
-  uint64_t l_tmp_1 = i_a[1];
-  uint64_t l_tmp_2 = i_a[2];
-  uint64_t l_tmp_3 = i_a[3];
-  uint64_t l_tmp_4 = i_a[4];
-  uint64_t l_tmp_5 = i_a[5];
-  uint64_t l_tmp_6 = i_a[6];
-
-  i_b[0] = l_tmp_0;
-  i_b[1] = l_tmp_1;
-  i_b[2] = l_tmp_2;
-  i_b[3] = l_tmp_3;
-  i_b[4] = l_tmp_4;
-  i_b[5] = l_tmp_5;
-  i_b[6] = l_tmp_6;
-}

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_22_12_12_dressler_kroehl_halank/src/copy_driver.cpp

@@ -1,56 +1,0 @@
-#include <cstdint>
-#include <cstdlib>
-#include <iostream>
-
-extern "C"
-{
-  void copy_c(uint64_t const *i_a,
-              uint64_t *o_b);
-
-  void copy_asm(uint64_t const *i_a, // x0
-                uint64_t *o_b);      // x1
-                
-  void copy_asm_loop(uint64_t const *i_a, //x0
-                     uint64_t *o_b);      //x1
-}
-
-int main()
-{
-  uint64_t l_a[7] = {1, 21, 43, 78, 89, 91, 93};
-
-  uint64_t l_b_0[7] = {0};
-  uint64_t l_b_1[7] = {0};
-  uint64_t l_b_2[7] = {0};
-
-  // copy_c
-  std::cout << "### calling copy_c ###" << std::endl;
-  copy_c(l_a,
-         l_b_0);
-
-  for (unsigned short l_va = 0; l_va < 7; l_va++)
-  {
-    std::cout << l_a[l_va] << " / " << l_b_0[l_va] << std::endl;
-  }
-
-  // copy_asm
-  std::cout << "### calling copy_asm ###" << std::endl;
-  copy_asm(l_a,
-           l_b_1);
-
-  for (unsigned short l_va = 0; l_va < 7; l_va++)
-  {
-    std::cout << l_a[l_va] << " / " << l_b_1[l_va] << std::endl;
-  }
-
-  // copy_asm_loop
-  std::cout << "### calling copy_asm_loop ###" << std::endl;
-  copy_asm_loop(l_a,
-           l_b_2);
-
-  for (unsigned short l_va = 0; l_va < 7; l_va++)
-  {
-    std::cout << l_a[l_va] << " / " << l_b_2[l_va] << std::endl;
-  }
-
-  return EXIT_SUCCESS;
-}

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_22_12_12_dressler_kroehl_halank/src/fibonacci_c.c

@@ -1,18 +1,0 @@
-#include <stdint.h>
-
-// returns n-th fibonacci number
-uint64_t fibonacci_c( uint64_t i_n ) {
-  if( i_n <= 1 ) return i_n;
-
-  uint64_t l_tmp_0 = 0;
-  uint64_t l_tmp_1 = 1;
-  uint64_t l_tmp_2 = 0;
-
-  for( uint64_t l_va = 1; l_va < i_n; l_va++ ) {
-    l_tmp_2 = l_tmp_0 + l_tmp_1;
-    l_tmp_0 = l_tmp_1;
-    l_tmp_1 = l_tmp_2;
-  }
-
-  return l_tmp_2;
-}

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_22_12_12_dressler_kroehl_halank/src/fibonacci_driver.cpp

@@ -1,48 +1,0 @@
-#include <cstdint>
-
-#include <cstdlib>
-
-#include <iostream>
-
-
-extern "C" {
-
-  uint64_t fibonacci_c( uint64_t i_id );
-
-  uint64_t fibonacci_asm( uint64_t i_id );
-
-}
-
-
-int main() {
-
-  uint64_t l_id = 50;
-
-  uint64_t l_number_0 = 0;
-
-  uint64_t l_number_1 = 0;
-
-
-  // fibonacci_c
-
-  std::cout << "### fibonacci_c ###" << std::endl;
-
-  l_number_0 = fibonacci_c( l_id );
-
-
-  std::cout << "id / number: " << l_id << " / " << l_number_0 << std::endl;
-
-
-  // fibonacci_asm
-
-  std::cout << "### fibonacci_asm ###" << std::endl;
-
-  l_number_1 = fibonacci_asm( l_id );
-
-
-  std::cout << "id / number: " << l_id << " / " << l_number_1 << std::endl;
-
-
-  return EXIT_SUCCESS;
-
-}

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_23_01_02_dressler_kroehl_halank/src/driver.cpp

@@ -1,230 +1,0 @@
-#include <iostream>
-
-#include "high_level.h"
-#include "low_level.h"
-
-int main()
-{
-
-    std::cout << "running driver" << std::endl;
-
-    std::cout << "high_lvl_0(10): "
-
-              << high_lvl_0(10)
-
-              << std::endl;
-
-    std::cout << "high_lvl_1(10): "
-
-              << high_lvl_1(10) << std::endl;
-
-    std::cout << "high_lvl_2(32): "
-
-              << high_lvl_2(32) << std::endl;
-
-    std::cout << "high_lvl_2( 5): "
-
-              << high_lvl_2(5) << std::endl;
-
-    int32_t l_highLvlOpt3 = 17;
-
-    int32_t l_highLvlRes3 = -1;
-
-    high_lvl_3(&l_highLvlOpt3,
-
-               &l_highLvlRes3);
-
-    std::cout << "high_lvl_3 #1: "
-
-              << l_highLvlRes3 << std::endl;
-
-    l_highLvlOpt3 = 43;
-
-    high_lvl_3(&l_highLvlOpt3,
-
-               &l_highLvlRes3);
-
-    std::cout << "high_lvl_3 #2: "
-
-              << l_highLvlRes3 << std::endl;
-
-    std::cout << "high_lvl_4(1,2,3): "
-
-              << high_lvl_4(1, 2, 3) << std::endl;
-
-    std::cout << "high_lvl_4(4,2,3): "
-
-              << high_lvl_4(4, 2, 3) << std::endl;
-
-    std::cout << "high_lvl_4(4,3,3): "
-
-              << high_lvl_4(4, 3, 3) << std::endl;
-
-    int32_t l_highLvlValue5 = 500;
-
-    high_lvl_5(17,
-
-               &l_highLvlValue5);
-
-    std::cout << "high_lvl_5: " << l_highLvlValue5 << std::endl;
-
-    int64_t l_highLvlValue6 = 23;
-
-    high_lvl_6(5,
-
-               13,
-
-               &l_highLvlValue6);
-
-    std::cout << "high_lvl_6: "
-
-              << l_highLvlValue6 << std::endl;
-
-    int64_t l_highLvlVasIn7[10] = {0, 7, 7, 4, 3,
-
-                                   -10, -50, 40, 2, 3};
-
-    int64_t l_highLvlVasOut7[10] = {0};
-
-    high_lvl_7(10,
-
[…]

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_23_01_02_dressler_kroehl_halank/src/high_level.cpp

@@ -1,120 +1,0 @@
-#include "high_level.h"
-
-int32_t high_lvl_0(int32_t i_value)
-{
-
-    return i_value;
-}
-
-uint64_t high_lvl_1(uint64_t)
-{
-
-    return 0;
-}
-
-int32_t high_lvl_2(int32_t i_option)
-{
-
-    int32_t l_result = 0;
-
-    if (i_option < 32)
-    {
-
-        l_result = 1;
-    }
-
-    return l_result;
-}
-
-void high_lvl_3(int32_t *i_option,
-
-                int32_t *o_result)
-{
-
-    if (*i_option < 25)
-    {
-
-        *o_result = 1;
-    }
-
-    else
-    {
-
-        *o_result = 0;
-    }
-}
-
-uint32_t high_lvl_4(uint32_t i_x,
-
-                    uint32_t i_y,
-
-                    uint32_t i_z)
-{
-
-    uint32_t l_ret = 0;
-
-    if (i_x < i_y && i_x < i_z)
-    {
-
-        l_ret = 1;
-    }
-
-    else if (i_y < i_z)
-    {
-
-        l_ret = 2;
-    }
-
-    else
-    {
-
-        l_ret = 3;
-    }
-
-    return l_ret;
-}
-
-void high_lvl_5(uint32_t i_nIters,
-
-                int32_t *io_value)
-{
-
-    for (uint32_t l_i = 0; l_i < i_nIters; l_i++)
-    {
-
-        *io_value += 1;
-    }
-}
-
-void high_lvl_6(uint64_t i_nIters,
-
-                int64_t i_inc,
-
-                int64_t *io_value)
-{
-
-    uint64_t l_va = i_nIters;
-
-    do
-    {
-
-        *io_value += i_inc;
-
-        l_va--;
-
-    } while (l_va != 0);
-}
-
-void high_lvl_7(uint64_t i_nValues,
-
-                int64_t *i_valuesIn,
-
-                int64_t *i_valuesOut)
-{
-
-    for (uint64_t l_va = 0; l_va < i_nValues; l_va++)
-    {
-
-        i_valuesOut[l_va] = i_valuesIn[l_va];
-    }
-}

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_23_01_02_dressler_kroehl_halank/src/high_level.h

@@ -1,41 +1,0 @@
-#include <cstdint>
-
-
-int32_t high_lvl_0( int32_t i_value );
-
-
-uint64_t high_lvl_1( uint64_t );
-
-
-int32_t high_lvl_2( int32_t i_option );
-
-
-void high_lvl_3( int32_t * i_option,
-
-                 int32_t * o_result );
-
-
-uint32_t high_lvl_4( uint32_t i_x,
-
-                     uint32_t i_y,
-
-                     uint32_t i_z );
-
-
-void high_lvl_5( uint32_t   i_nIters,
-
-                 int32_t  * io_value );
-
-
-void high_lvl_6( uint64_t   i_nIters,
-
-                 int64_t    i_inc,
-
-                 int64_t  * io_value );
-
-
-void high_lvl_7( uint64_t   i_nValues,
-
-                 int64_t  * i_valuesIn,
-
-                 int64_t  * i_valuesOut );

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_23_01_02_dressler_kroehl_halank/src/low_level.h

@@ -1,36 +1,0 @@
-#include <cstdint>
-
-extern "C"
-{
-    int32_t low_lvl_0(int32_t i_value);
-
-    uint64_t low_lvl_1(uint64_t);
-
-    int32_t low_lvl_2(int32_t i_option);
-
-    void low_lvl_3(int32_t *i_option,
-
-                   int32_t *o_result);
-
-    uint32_t low_lvl_4(uint32_t i_x,
-
-                       uint32_t i_y,
-
-                       uint32_t i_z);
-
-    void low_lvl_5(uint32_t i_nIters,
-
-                   int32_t *io_value);
-
-    void low_lvl_6(uint64_t i_nIters,
-
-                   int64_t i_inc,
-
-                   int64_t *io_value);
-
-    void low_lvl_7(uint64_t i_nValues,
-
-                   int64_t *i_valuesIn,
-
-                   int64_t *i_valuesOut);
-}

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_23_01_09_dressler_kroehl_halank/src/9_1/driver.cpp

@@ -1,24 +1,0 @@
-#include <cstdint>
-#include <cstdlib>
-#include <iostream>
-
-extern "C" {
-    void showcase_fmla_element( float const * i_a,
-                            float const * i_b,
-                            float       * o_c);
-}
-
-int main(){
-    std::cout << "Running 9.1.2 Code" << std::endl;
-
-    float l_a_fp32[4] = {1,2,3,4};
-    float l_b_fp32[4] = {5,6,7,8};
-    float l_c_fp32[4] = {10,11,12,13};
-
-    showcase_fmla_element(l_a_fp32, l_b_fp32, l_c_fp32);
-
-    std::cout << "C:" << l_c_fp32[0] << " " << l_c_fp32[1] << " " << l_c_fp32[2] << " " << l_c_fp32[3] << std::endl;
-    std::cout << "Finished 9.1.2 Code" << std::endl;
-    return EXIT_SUCCESS;
-}
-    

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_23_01_09_dressler_kroehl_halank/src/9_2/driver.cpp

@@ -1,191 +1,0 @@
-#include <iostream>
-#include <cmath>
-#include <chrono>
-#include "kernels/gemm_ref.h"
-
-const unsigned int l_n_repetitions = 100000000;
-
-extern "C" {
-  void gemm_asm_fp_4_4_4( float const * i_a,
-                          float const * i_b,
-                          float       * io_c );
-  void gemm_asm_asimd_4_4_4( float const * i_a,
-                             float const * i_b,
-                             float       * io_c );
-}
-
-/**
- * Computes the maximum absolute element-wise difference of matrices.
- *
- * @param i_mat0 first matrix (column-major).
- * @param i_mat1 second matrix (column-major).
- * @param i_m number of rows.
- * @param i_n number of columns.
- * @param i_ld leading dimension.
- **/
-float max_diff( float const * i_mat0,
-                float const * i_mat1,
-                unsigned int  i_m,
-                unsigned int  i_n,
-                unsigned int  i_ld ) {
-  float l_max_diff = 0;
-
-  for( unsigned int l_m = 0; l_m < i_m; l_m++ ) {
-    for( unsigned int l_n = 0; l_n < i_n; l_n++ ) {
-      float l_diff = i_mat0[ l_n*i_ld + l_m ] - i_mat1[ l_n*i_ld + l_m ];
-      l_diff = std::abs( l_diff );
-
-      l_max_diff = std::max( l_max_diff, l_diff );
-    }
-  }
-
-  return l_max_diff;
-}
-
-int main() {
-  // allocate memory
-  std::size_t l_size = 64*64;
-  float * l_a = new float[ l_size ];
-  float * l_b = new float[ l_size ];
-  float * l_c = new float[ l_size ];
-  float * l_c_ref = new float[ l_size ];
-
-  // init seed
-  srand48( time(NULL) );
-
-  std::chrono::steady_clock::time_point l_tp0, l_tp1;
-  std::chrono::duration<double> l_dur;
-  float l_max_diff = 0;
-  double l_g_flops =0;
-
-  /*
-   * FP 4, 4, 4
-   */
-  std::cout << "testing gemm_asm_fp_4_4_4 kernel" << std::endl;
-
-  // reset data
-  for( unsigned int l_id = 0; l_id < l_size; l_id++ ) {
-    l_a[l_id] = (float) drand48();
-  }
[…]

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_23_01_09_dressler_kroehl_halank/src/9_2/kernels/gemm_ref.cpp

@@ -1,17 +1,0 @@
-void gemm_ref_mnk( float const  * i_a,
-                   float const  * i_b,
-                   float        * io_c,
-                   unsigned int   i_m,
-                   unsigned int   i_n,
-                   unsigned int   i_k,
-                   unsigned int   i_lda,
-                   unsigned int   i_ldb,
-                   unsigned int   i_ldc ) {
-  for( unsigned int l_m = 0; l_m < i_m; l_m++ ) {
-    for( unsigned int l_n = 0; l_n < i_n; l_n++ ) {
-      for( unsigned int l_k = 0; l_k < i_k; l_k++ ) {
-        io_c[ l_n*i_ldc + l_m ] += i_a[ l_k*i_lda + l_m ] * i_b[ l_n*i_ldb + l_k ];
-      }
-    }
-  }
-}

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_23_01_09_dressler_kroehl_halank/src/9_2/kernels/gemm_ref.h

@@ -1,22 +1,0 @@
-/**
- * Reference implementation of a GEMM (C+=AB).
- *
- * @param i_a pointer to A (column-major).
- * @param i_b pointer to B (column-major)
- * @param io_c pointer to C (column-major).
- * @param i_m BLAS parameter M.
- * @param i_n BLAS parameter N.
- * @param i_k BLAS parameter K.
- * @param i_lda leading dimension of A.
- * @param i_ldb leading dimension of B.
- * @param i_ldc leading dimension of C.
- **/
-void gemm_ref_mnk( float const  * i_a,
-                   float const  * i_b,
-                   float        * io_c,
-                   unsigned int   i_m,
-                   unsigned int   i_n,
-                   unsigned int   i_k,
-                   unsigned int   i_lda,
-                   unsigned int   i_ldb,
-                   unsigned int   i_ldc );

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_23_01_16_dressler_kroehl_halank/src/asm_driver.cpp

@@ -1,32 +1,0 @@
-#include <cstdint>
-#include <cstdlib>
-#include <iostream>
-
-extern "C" {
-  uint64_t machine_code_c();
-  uint64_t machine_code_asm_0();
-  uint64_t machine_code_asm_1();
-}
-
-int main() {
-  uint64_t l_result_c = 0;
-  uint64_t l_result_asm_0 = 0;
-  uint64_t l_result_asm_1 = 0;
-
-  // machine_code_c
-  std::cout << "### calling machine_code_c ###" << std::endl;
-  l_result_c = machine_code_c();
-  std::cout << l_result_c << std::endl;
-
-  // machine_code_asm_0
-  std::cout << "### calling machine_code_asm_0 ###" << std::endl;
-  l_result_asm_0 = machine_code_asm_0();
-  std::cout << l_result_asm_0 << std::endl;
-
-  // machine_code_asm_1
-  std::cout << "### calling machine_code_asm_1 ###" << std::endl;
-  l_result_asm_1 = machine_code_asm_1();
-  std::cout << l_result_asm_1 << std::endl;
-
-  return EXIT_SUCCESS;
-}

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_23_01_16_dressler_kroehl_halank/src/asm_kernel.c

@@ -1,19 +1,0 @@
-#include <stdint.h>
-
-uint64_t machine_code_c() {
-  uint64_t l_tmp_0 = 3;
-  uint64_t l_tmp_1 = 0;
-  uint64_t l_tmp_2 = 0;
-
-  while( 1 ) {
-    l_tmp_0--;
-    l_tmp_1 = l_tmp_1 + 3;
-    l_tmp_2 = l_tmp_2 + 7;
-
-    if( l_tmp_0 == 0 ) break;
-  }
-
-  l_tmp_0 = l_tmp_1 + l_tmp_2;
-
-  return l_tmp_0;
-}

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_23_01_30_dressler_kroehl_halank/src/micro_bench/driver_aarch64.cpp

@@ -1,108 +1,0 @@
-#include <cstdint>
-#include <cmath>
-#include <chrono>
-#include <tuple>
-#include <iostream>
-#include <unistd.h>
-
-extern "C" {
-  uint64_t micro_aarch64_madd_w_independent( uint64_t i_n_repetitions );
-  uint64_t micro_aarch64_madd_w_raw_rn( uint64_t i_n_repetitions );
-}
-
-/**
- * Measures the performance (time) of the given kernel.
- *
- * @param i_kernel benchmarked kernel.
- * @param i_n_repetitions number of performed repetitions.
- * @return duration required for the performed repetitions.
- **/
-double timeKernel( uint64_t (*i_kernel)(uint64_t),
-                   uint64_t   i_n_repetitions ) {
-  std::chrono::steady_clock::time_point l_tp0, l_tp1;
-  std::chrono::duration< double > l_dur;
-
-  l_tp0 = std::chrono::steady_clock::now();
-  (*i_kernel)( i_n_repetitions );
-  l_tp1 = std::chrono::steady_clock::now();
-
-  l_dur = std::chrono::duration_cast< std::chrono::duration< double> >( l_tp1 - l_tp0 );
-
-  return l_dur.count();
-}
-
-/**
- * Benchmarks the performance (repetitions, time, giga instructions) of the given microkernel.
- *
- * @param i_kernel benchmarked kernel.
- * @param i_time_target targeted total execution time; the number of actual repetitions is adjusted accordingly.
- * @param i_n_repetitions_initial initial number of performed repetitions.
- * @return (repetitions, time, gints).
- **/
-std::tuple< uint64_t,
-            double,
-            double > benchKernel( uint64_t (*i_kernel)(uint64_t),
-                                  double     i_time_target = 10.0,
-                                  uint64_t   i_n_repetitions_initial = 1000000 ) {
-  // get number of flops per iter
-  uint64_t l_n_insts = i_kernel(1);
-
-  // get time required for initial number of reps
-  double l_dur = timeKernel( i_kernel,
-                             i_n_repetitions_initial );
-
-  // derive number of reps for targeted duration
-  double l_scaling_time = i_time_target / l_dur;
-  uint64_t l_n_repetitions_adj = i_n_repetitions_initial * l_scaling_time;
[…]

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_23_02_06_dressler_kroehl_halank/src/sum.cpp

@@ -1,75 +1,0 @@
-#include <iostream>
-#include <chrono>
-#include <fstream>
-#include <iostream>
-using namespace std;
-
-int main()
-{
-  std::cout << "running benchmark" << std::endl;
-
-  ofstream myfile;
-  myfile.open ("output.txt");
-  myfile << "size\tbandwidth\n";
-
-  for (uint64_t l_size = 512; l_size <= 1024 * 1024 * 1024 / 4; l_size *= 2)
-  {
-
-    // derive size in bytes and gigabytes
-    uint64_t l_size_b = l_size * sizeof(uint64_t);
-    double l_size_gib = l_size_b / (1024.0 * 1024.0 * 1024.0);
-    uint64_t l_n_repetitions = (uint64_t)10 * 1024 * 1024 * 1024 / l_size_b;
-
-    // share config
-    std::cout << "config:\n"
-              << "  number of repetitions:        " << l_n_repetitions << "\n"
-              << "  number of entries per vector: " << l_size << "\n"
-              << "  size per vector (B):          " << l_size_b << "\n"
-              << "  size per vector (GB):         " << l_size_gib << "\n"
-              << "  total size (GB):              " << l_n_repetitions * l_size_gib
-              << std::endl;
-
-    // instrumentation vars
-    std::chrono::steady_clock::time_point l_tp0, l_tp1;
-    std::chrono::duration<double> l_dur;
-
-    // alloc memory and init
-    uint64_t *l_a = (uint64_t *)aligned_alloc(4096,
-                                              l_size_b);
-
-    for (uint64_t l_en = 0; l_en < l_size; l_en++)
-    {
-      l_a[l_en] = 1;
-    }
-
-    // benchmark
-    std::cout << "benchmarking" << std::endl;
-    uint64_t l_sum = 0;
-    l_tp0 = std::chrono::steady_clock::now();
-    for (uint64_t l_re = 0; l_re < l_n_repetitions; l_re++)
-    {
-      for (uint64_t l_en = 0; l_en < l_size; l_en++)
-      {
-        l_sum += l_a[l_en];
-      }
-    }
-    l_tp1 = std::chrono::steady_clock::now();
-
-    // share results
-    l_dur = std::chrono::duration_cast<std::chrono::duration<double>>(l_tp1 - l_tp0);
-    double l_bandwidth = l_n_repetitions * l_size_gib / l_dur.count();
[…]

### 2024-04-09 Archive/WiSe_2022_2023/TechnischeInformatik/submissions/submission_23_02_06_dressler_kroehl_halank/src/vector_product.cpp

@@ -1,62 +1,0 @@
-#include <cstdint>
-#include <cmath>
-#include <chrono>
-#include <tuple>
-#include <iostream>
-#include <unistd.h>
-
-uint64_t read(const uint64_t *a, const uint64_t *b, const uint64_t size, uint64_t distance)
-{
-    uint64_t result = 0;
-    for (uint64_t l = 0; l < distance; l++)
-    {
-        for (uint64_t k = 0; k < size / distance; k++)
-        {
-            result += a[distance * k + l] * b[distance * k + l];
-        }
-    }
-    return result;
-}
-
-int main()
-{
-    uint64_t l_size = 134217728;
-
-    // derive size in bytes and gigabytes
-    uint64_t l_size_b = l_size * sizeof(uint64_t);
-    double l_size_gib = l_size_b / (1024.0 * 1024.0 * 1024.0);
-
-    uint64_t *a = (uint64_t *)aligned_alloc(4096, l_size_b);
-    uint64_t *b = (uint64_t *)aligned_alloc(4096, l_size_b);
-
-    for (uint64_t l_en = 0; l_en < l_size; l_en++)
-    {
-        a[l_en] = 1;
-        b[l_en] = 2;
-    }
-
-    std::chrono::steady_clock::time_point start, end;
-    std::chrono::duration<double> l_dur;
-
-    for (int distance = 1; distance <= 16; distance *= 2)
-    {
-        // benchmark
-        start = std::chrono::steady_clock::now();
-        uint64_t result = read(a, b, l_size, distance);
-        end = std::chrono::steady_clock::now();
-
-        // share results
-        l_dur = std::chrono::duration_cast<std::chrono::duration<double>>(end - start);
-        std::cout << "Dot product Distance " << distance << ": \n"
-                  << "  error:             " << result - 2 * l_size << "\n"
-                  << "  time:              " << l_dur.count() << "\n"
-                  << "  bandwidth (GiB/s): " << l_size_gib / l_dur.count()
-                  << std::endl;
-    }
-
-    // free memory
-    free(a);
-    free(b);
-
-    return EXIT_SUCCESS;
-}

### 2024-04-10 PythonASQ/1/01_skript.py
Python Übung 1

@@ -1,0 +1,3 @@
+print('Hallo! Ich bin (wahrscheinlich) das erste Pythonskript welches du direkt auf der Kommandozeile ausgeführt hast!')    #Und ich bin ein Kommentar, ich kann direkt hinter Pythoncode stehen.
+
+# Ich bin auch ein Kommentar, ich kann auch auf einer eigenen Zeile stehen.

### 2024-04-10 PythonASQ/1/3.py

@@ -1,0 +1,9 @@
+x=y=1
+x=2
+z=2
+z=2*3
+x,y=1,6
+y=6/2
+z,x=3.0,6
+x="y"
+print(x)

### 2024-06-19 PythonASQ/11/mergeAnnotation.py
7GB implementation of merge Annotation.py

@@ -1,0 +1,60 @@
+from pathlib import Path
+import sys
+import resource
+import gc
+
+def chrAndStart(str: list[str]):
+    x = str.split("\t")
+    # get chromosome number
+    chr = int(x[0][3:])
+    # get start position
+    start = int(x[1])
+    return chr, start
+
+if len(sys.argv) != 3:
+    print("Usage: python mergeAnnotation.py <file1> <file2>")
+    sys.exit(1)
+
+file1 = sys.argv[1]
+file2 = sys.argv[2]
+usedMemory = 0
+sharedList = []
+
+with open(file1, "r") as f:
+    for i, line in enumerate(f):
+        if i == 0:
+            continue
+        sharedList.append(line)
+    
+with open(file2, "r") as f:
+    for i, line in enumerate(f):
+        if i == 0:
+            continue
+        sharedList.append(line)
+
+sharedList.sort(key=chrAndStart)
+
+# merge overlapping regions with same chromosome number
+
+mergedList = []
+mergedList.append(sharedList[0])
+
+for i in range(1, len(sharedList)):
+    shared_list_i = sharedList[i].split()
+    merged_list_min_1 = mergedList[-1].split()
+    if shared_list_i[0] == merged_list_min_1[0] and shared_list_i[1] <= merged_list_min_1[2]:
+        mergedList[-1] = f"{merged_list_min_1[0]}\t{merged_list_min_1[1]}\t{shared_list_i[2]}\t{merged_list_min_1[3]}{shared_list_i[3]}\n"
+    else:
+        mergedList.append(sharedList[i])
+
+del sharedList
+gc.collect()
+
+Path("bin").mkdir(parents=True, exist_ok=True)
+with open("bin/mergedAnnotation.bed", "w") as f:
+    f.write("chr\tstart\tend\tannotation\n")
+    for line in mergedList:
+        f.write(line)
+
+statPeakMemoryMB = resource.getrusage(resource.RUSAGE_SELF).ru_maxrss / 1024
+print(f"Peak memory usage: {statPeakMemoryMB:.2f} MB")

### 2024-06-19 PythonASQ/11/sorting.py

@@ -98,34 +98,35 @@ def quickSort(A: list, lo = 0, hi = None):
     
     return A
 
-n = 1000
-testList = randList.make_random_list_choices(n)
-
-testListInsertion = testList.copy()
-testListInsertionBinary = testList.copy()
-testListSelection = testList.copy()
-testListMergeSort = testList.copy()
-testListQuickSort = testList.copy()
-
-laufzeitInsertion = timeit.timeit(lambda: insertionSort(testListInsertion), globals=globals(), number=1)
-laufzeitInsertionBinary = timeit.timeit(lambda: binaryInsertionSort(testListInsertionBinary), globals=globals(), number=1)
-laufzeitSelection = timeit.timeit(lambda: selectionSort(testListSelection), globals=globals(), number=1)
-laufzeitQuickSort = timeit.timeit(lambda: quickSort(testListQuickSort), globals=globals(), number=1)
-laufzeitMergeSort = timeit.timeit(lambda: mergeSort(testListMergeSort), globals=globals(), number=1)
-laufzeitPowerSort = timeit.timeit(lambda: testList.sort(), globals=globals(), number=1)
-
-print(f"InsertionSort: {laufzeitInsertion}s")
-print(f"BinaryInsertionSort: {laufzeitInsertionBinary}s")
-print(f"SelectionSort: {laufzeitSelection}s")
-print(f"QuickSort: {laufzeitQuickSort}s")
-print(f"MergeSort: {laufzeitMergeSort}s")
-print(f"PowerSort: {laufzeitPowerSort}s") # assumes at least Python 3.11
-
-"""
-PowerSort << Quicksort < MergeSort << BinaryInsertionSort << InsertionSort < SelectionSort
-"""
-
-# profile.run("insertionSort(testListInsertion)", sort="ncalls")
-# profile.run("binaryInsertionSort(testListInsertionBinary)", sort="ncalls")
-# profile.run("selectionSort(testListSelection)", sort="ncalls")
+if __name__ == "__main__":
+    n = 1000
+    testList = randList.make_random_list_choices(n)
+    
+    testListInsertion = testList.copy()
+    testListInsertionBinary = testList.copy()
+    testListSelection = testList.copy()
+    testListMergeSort = testList.copy()
+    testListQuickSort = testList.copy()
+    
+    laufzeitInsertion = timeit.timeit(lambda: insertionSort(testListInsertion), globals=globals(), number=1)
[…]

### 2024-06-19 PythonASQ/11/appendAndSortList.cpp
Added Low Memory Merge (its CPP ok?)

@@ -1,0 +1,78 @@
+#include <iostream>
+#include <fstream>
+#include <vector>
+#include <string>
+#include <sys/resource.h>
+
+bool compare(const std::string &a, const std::string &b)
+{
+    std::string delimiter = "\t";
+    size_t posA1 = a.find(delimiter);
+    size_t posB1 = b.find(delimiter);
+    int aChr = std::stoi(a.substr(3, posA1));
+    int bChr = std::stoi(b.substr(3, posB1));
+    if (aChr != bChr)
+    {
+        return aChr < bChr;
+    }
+    size_t posA2 = a.find(delimiter, posA1 + 1);
+    size_t posB2 = b.find(delimiter, posB1 + 1);
+    std::string aPos = a.substr(posA1 + 1, posA2);
+    std::string bPos = b.substr(posB1 + 1, posB2);
+    return aPos.length() == bPos.length() && aPos < bPos || aPos.length() < bPos.length();
+}
+
+int main(int argc, char const *argv[])
+{
+    if (argc != 3)
+    {
+        std::cerr << "Usage: " << argv[0] << " <file1> <file2>" << std::endl;
+        return 1;
+    }
+    std::string file1 = argv[1];
+    std::string file2 = argv[2];
+
+    std::ifstream file1Stream(file1);
+    std::ifstream file2Stream(file2);
+
+    if (!file1Stream.is_open())
+    {
+        std::cerr << "Error: Could not open file " << file1 << std::endl;
+        return 1;
+    }
+    if (!file2Stream.is_open())
+    {
+        std::cerr << "Error: Could not open file " << file2 << std::endl;
+        return 1;
+    }
+
+    std::vector<std::string> shared_list;
+    std::string line;
+    getline(file1Stream, line);
+    while (getline(file1Stream, line))
+    {
+        shared_list.push_back(line);
+    }
+    getline(file2Stream, line);
+    while (getline(file2Stream, line))
+    {
+        shared_list.push_back(line);
+    }
+
+    std::sort(shared_list.begin(), shared_list.end(), compare);
+
+    std::ofstream out("bin/sorted.bed");
+    for (auto &line : shared_list)
+    {
+        out << line << std::endl;
+    }
+
+    struct rusage usage;
+    getrusage(RUSAGE_SELF, &usage);
+    auto statPeakMemoryMB = usage.ru_maxrss / 1024;
+
[…]

### 2024-06-19 PythonASQ/11/mergeAnnotation.py

@@ -17,7 +17,6 @@ if len(sys.argv) != 3:
 
 file1 = sys.argv[1]
 file2 = sys.argv[2]
-usedMemory = 0
 sharedList = []
 
 with open(file1, "r") as f:
@@ -42,8 +41,8 @@ mergedList.append(sharedList[0])
 for i in range(1, len(sharedList)):
     shared_list_i = sharedList[i].split()
     merged_list_min_1 = mergedList[-1].split()
-    if shared_list_i[0] == merged_list_min_1[0] and shared_list_i[1] <= merged_list_min_1[2]:
-        mergedList[-1] = f"{merged_list_min_1[0]}\t{merged_list_min_1[1]}\t{shared_list_i[2]}\t{merged_list_min_1[3]}{shared_list_i[3]}\n"
+    if shared_list_i[0] == merged_list_min_1[0] and int(shared_list_i[1]) <= int(merged_list_min_1[2]):
+        mergedList[-1] = f"{merged_list_min_1[0]}\t{merged_list_min_1[1]}\t{max(int(shared_list_i[2]),int(merged_list_min_1[2]))}\t{merged_list_min_1[3]}{shared_list_i[3]}\n"
     else:
         mergedList.append(sharedList[i])
 

### 2024-06-19 PythonASQ/11/mergeAnnotationLowMem.py

@@ -1,0 +1,40 @@
+from pathlib import Path
+import sys
+import resource
+import subprocess
+
+if len(sys.argv) != 3:
+    print("Usage: python mergeAnnotationLowMem.py <file1> <file2>")
+    sys.exit(1)
+
+file1 = sys.argv[1]
+file2 = sys.argv[2]
+
+# check if appendAndSortList.out exists
+if not Path("bin/appendAndSort.out").is_file():
+    subprocess.run(["g++","-std=c++20", "appendAndSortList.cpp", "-O3","-march=native", "-o", "bin/appendAndSort.out"])    
+
+subprocess.run(["bin/appendAndSort.out", file1, file2])
+# merge overlapping regions with same chromosome number
+
+mergedList = []
+with open("bin/sorted.bed", "r") as f:
+    for i, line in enumerate(f):
+        if i == 0:
+            mergedList.append(line)
+            continue
+        shared_list_i = line.split()
+        merged_list_min_1 = mergedList[-1].split()
+        if shared_list_i[0] == merged_list_min_1[0] and int(shared_list_i[1]) <= int(merged_list_min_1[2]):
+            mergedList[-1] = f"{merged_list_min_1[0]}\t{merged_list_min_1[1]}\t{max(int(shared_list_i[2]),int(merged_list_min_1[2]))}\t{merged_list_min_1[3]}{shared_list_i[3]}\n"
+        else:
+            mergedList.append(line)
+
+Path("bin").mkdir(parents=True, exist_ok=True)
+with open("bin/mergedAnnotation.bed", "w") as f:
+    f.write("chr\tstart\tend\tannotation\n")
+    for line in mergedList:
+        f.write(line)
+
+statPeakMemoryMB = resource.getrusage(resource.RUSAGE_SELF).ru_maxrss / 1024
+print(f"Peak Memory Python: {statPeakMemoryMB:.2f} MB")

### 2024-06-25 PythonASQ/12/binaryTree.py
Added BinaryTrees

@@ -1,0 +1,572 @@
+import math
+import random
+
+
+class Graph:
+    class Node:
+        def __init__(self, name, value = 0):
+            self._name = name
+            self._edges: list[Graph.Node] = []
+            self.value = value
+                    
+        @property
+        def name(self):
+            return self._name
+        
+        @property
+        def edges(self):
+            return self._edges
+        
+        def deg(self):
+            return len(self.edges)
+        
+        def __repr__(self):
+            return f"Node<{self.name}>"
+        
+    def __init__(self, nodes : list[str] = None):
+        self.nodes: dict[str,Graph.Node] = {}
+        if nodes is not None:
+            if len(nodes) != len(set(nodes)):
+                raise ValueError('Knoten doppelt vorhanden')
+            for node in nodes:
+                self.add_node(node)
+        self.num_edges = 0
+        self.edge_weights = {}
+    
+    @property
+    def num_nodes(self):
+        return len(self.nodes)
+    
+    def add_node(self, name, value = 0):
+        if name in self.nodes:
+            raise ValueError('Knoten bereits vorhanden')
+        self.nodes[name] = Graph.Node(name, value)
+        return self.nodes[name]
+    
+    def change_name(self, oldName, newName):
+        if newName in self.nodes:
+            raise ValueError('neuer Name bereits vorhanden')
+        try:
+            node = self.nodes[oldName]
+        except:
+            raise ValueError('Knoten nicht vorhanden')
+        del self.nodes[oldName]
+        node._name = newName
+        self.nodes[newName] = node
+    
+    def remove_node(self, name):
+        try:
+            node = self.nodes[name]
+        except KeyError:
+            raise ValueError('Knoten nicht vorhanden')
+        del self.nodes[name]
+        for n in node.edges:
+            n._edges.remove(node)
+        self.num_edges -= node.deg()
+        
+    def add_edge(self, name1: str, name2: str, value = 0):
+        try:
+            node1 = self.nodes[name1]
+            node2 = self.nodes[name2]
+        except KeyError:
+            raise ValueError('ein Knoten nicht vorhanden')
+        if node2 in node1.edges:
+            raise ValueError('Kante bereits vorhanden')
+        node1._edges.append(node2)
+        node2._edges.append(node1)
+        self.edge_weights[(name1, name2)] = value
+        self.edge_weights[(name2, name1)] = value
+        self.num_edges += 1
+        
+    def remove_edge(self, name1: str, name2: str):
+        try:
+            node1 = self.nodes[name1]
[…]

### 2024-06-25 PythonASQ/12/binaryTree.py
Implemented Dijkstra

@@ -183,7 +183,7 @@ class Graph:
             return True
         raise ValueError('key must be a string, a Graph.Node or a list of edges')
     
-    def breadthSearch(self, value: int ,startNode: Node = None) -> Node:
+    def bfs(self, value: int ,startNode: Node = None) -> Node:
         if startNode is None:
             startNode = self.nodes[next(iter(self.nodes))]
         queue = [startNode]
@@ -197,6 +197,32 @@ class Graph:
                 if edge.name not in visitedNames:
                     queue.append(edge)
         return None
+    
+    def dfs(self, value: int ,startNode: Node = None) -> Node:
+        if startNode is None:
+            startNode = self.nodes[next(iter(self.nodes))]
+        queue = [startNode]
+        visitedNames = set()
+        while queue:
+            node = queue.pop()
+            if node.value == value:
+                return node
+            visitedNames.add(node.name)
+            for edge in node.edges:
+                if edge.name not in visitedNames:
+                    queue.append(edge)
+        return None
+
+    def isConnected(self):
+        visited = set()
+        queue = [next(iter(self.nodes))]
+        while queue:
+            node = queue.pop()
+            visited.add(node)
+            for edge in node.edges:
+                if edge not in visited:
+                    queue.append(edge)
+        return len(visited) == len(self.nodes)
 
 # Meckerbemerkung: es wäre viel sinnvoller gewesen einen Bidirektionalen Graphen
 # als Spezialfall eines Gerichteten Graphen zu implementieren
@@ -415,11 +441,15 @@ def generateGraph(n: int, i: int, j: int)-> Graph:
     return graph
 
 def greedySearch(graph: Graph, start: str, end: str):
+    if start not in graph or end not in graph:
+        raise ValueError('Knoten nicht vorhanden')
+    start = graph[start]
+    end = graph[end]
     nodeOrder = [start]
     distance = []
     visited = set()
     while nodeOrder[-1] != end:
-        node = graph[nodeOrder[-1]]
+        node = nodeOrder[-1]
         visited.add(node.name)
         nextNode = None
         minWeight = float('inf')
@@ -432,9 +462,39 @@ def greedySearch(graph: Graph, start: str, end: str):
             nodeOrder.pop()
             distance.pop()
         else:
-            nodeOrder.append(nextNode.name)
+            nodeOrder.append(nextNode)
             distance.append(minWeight)
[…]

### 2024-06-25 PythonASQ/12/binaryTree.py
verbesserte Graphen generation

@@ -215,7 +215,9 @@ class Graph:
 
     def isConnected(self):
         visited = set()
-        queue = [next(iter(self.nodes))]
+        if not self.nodes:
+            raise ValueError('Graph is empty')
+        queue = [list(self.nodes.values())[0]]
         while queue:
             node = queue.pop()
             visited.add(node)
@@ -432,12 +434,15 @@ def generateGraph(n: int, i: int, j: int)-> Graph:
     graph = Graph()
     for k in range(n):
         graph.add_node(str(k), random.randint(i,j))
+    allNodes = set(graph.nodes)
     for node in graph.nodes.values():
-        # at least n/2 edges ensures a connected graph
-        for _ in range(random.randint(math.ceil(n/2),n)):
-            node2 = random.choice(list(graph.nodes.values()))
-            if node2 != node and node2 not in node.edges:
-                graph.add_edge(node.name, node2.name, random.randint(i,j))
+        currentEdges = set(map(lambda n: n.name, node.edges))
+        nEdges = len(currentEdges)
+        edgesToCreate = math.ceil(n/2)-nEdges
+        possibleEdges = sorted(allNodes - currentEdges - {node.name})
+        newEdges = random.sample(possibleEdges, edgesToCreate)
+        for edge in newEdges:
+            graph.add_edge(node.name, edge, random.randint(i,j))
     return graph
 
 def greedySearch(graph: Graph, start: str, end: str):
@@ -608,26 +613,45 @@ class Binärbaum(Graph):
         self.remove_node(node.name)
     
     
-graph = generateGraph(4, 1, 4)
-for node in graph.nodes.values():
-    print(node.name, list(map(lambda n: (n.name, graph.get_edge_weight(node.name,n.name)),node.edges)))
-print(graph.bfs(3))
-print(graph.dfs(3))
-print(greedySearch(graph, '0', '3'))
-print(dijkstra(graph,'0','3'))
+# graph = generateGraph(4, 1, 4)
+# for node in graph.nodes.values():
+#     print(node.name, list(map(lambda n: (n.name, graph.get_edge_weight(node.name,n.name)),node.edges)))
+# print(graph.bfs(3))
+# print(graph.dfs(3))
+# print(greedySearch(graph, '0', '3'))
+# print(dijkstra(graph,'0','3'))
 
-baum = Binärbaum()
-baum.build([0,1,2,3,4,5,6])
-print(baum.search(1))
[…]

### 2024-07-07 HPC/paper.tex
Python Übung 1
Added ToDo's to paper
Added BinaryTrees
Added Sections

@@
-%\documentclass[sigconf, language=german]{acmart}
-\documentclass[sigconf]{acmart} %TODO: Choose German or English
+\documentclass[sigconf]{acmart} %\documentclass[sigconf, language=german]{acmart}
 \usepackage{booktabs}
@@
 
-\title{MPI+OpenMP} %TODO: Add Title
+\title{MPI+OpenMP in MPI4.0}
 \author{Dreßler Justus}
@@
 
-% TODO: insert CCSXML http://dl.acm.org/ccs.cfm
+\begin{CCSXML}
+    <ccs2012>
+    <concept>
+    <concept_id>10010147.10010169</concept_id>
+    <concept_desc>Computing methodologies~Parallel computing methodologies</concept_desc>
+    <concept_significance>500</concept_significance>
+    </concept>
+    </ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Computing methodologies~Parallel computing methodologies}
 
@@
 
-%TODO: add content (preferibly in a new tex document)
+% Introduction, roughly 1 page
+\section{Introduction}
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Mauris dictum quam eget orci porttitor ultrices.
-Suspendisse purus nisl, dapibus et rhoncus in, eleifend id quam.
-Integer volutpat dolor urna, non auctor magna sodales non.
-Ut in fermentum erat, lacinia malesuada urna.
-Nullam et dolor ut purus fermentum rutrum.
-Suspendisse non velit sollicitudin, rutrum est at, commodo sem.
-Pellentesque viverra tincidunt mauris ut egestas.
-Integer vel orci sapien.
-Sed ut mi at lacus vehicula fermentum.
-Integer euismod vestibulum auctor.
-Nunc vulputate leo eget aliquam bibendum.
+Supercomputing relies on multiple layers of parallelisation, one intranode and one internode.
+Historically this was handled fine with a MPI+MPI (or MPI-Everywhere) programming model.
+In modern supercomputing rising core counts neccessitates a more nuanced programming model, one that makes scaling across cores on a node easier.
+This caused MPI+OpenMP to become the new de facto standard programming model in Supercomputing.
+It simplifies the seperation of internode and intranode tasks greatly, but it also imposes new challenges for developers.
+This paper tries to summarize the current (MPI4.0) methods to overcome these challenges and unlock high performance MPI+OpenMP applications.
+% this feels like making a buzzword list
 
-Etiam at cursus velit.
-Integer pellentesque, magna nec aliquet lobortis, sapien elit tempus ante, at eleifend eros risus sed dui.
-Suspendisse consectetur sed purus fermentum sollicitudin.
-Nam id sem et augue cursus luctus.
-Donec mauris neque, bibendum pellentesque fermentum sit amet, malesuada et augue.
-In ante enim, aliquam ac scelerisque id, blandit non massa.
-Donec placerat gravida ipsum et mattis.
-Cras scelerisque arcu in vulputate vehicula.
+\section{Motivation}
+% write why parallel communications is such a large burdon
 
-Donec nunc tortor, dignissim vitae ultricies et, mollis at turpis.
-Ut bibendum sit amet magna rutrum condimentum.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia curae; Morbi a vulputate velit.
-Suspendisse tellus est, ullamcorper in erat quis, iaculis aliquam diam.
-Aliquam pretium lobortis sollicitudin.
-In ac massa dictum, semper nisl id, condimentum ligula.
-Cras accumsan magna sed malesuada faucibus.
-Nulla in orci hendrerit, convallis tortor sed, vehicula nunc.
-Donec id magna bibendum, sodales purus non, congue neque.
-Quisque metus leo, lobortis ut metus a, dignissim lobortis felis.
+% Body, roughly 2.5 pages
+\section{Parallel Communication in MPI4.0}
 
-Fusce quis posuere metus, ac consequat diam.
-Donec sed venenatis nisl, in ultricies massa.
-Suspendisse feugiat vestibulum aliquam.
-Cras consequat venenatis feugiat.
-Morbi fermentum faucibus mauris eu rutrum.
-Sed eget scelerisque sem, vel volutpat dui.
-Nulla a laoreet arcu, vitae varius erat.
-Proin varius fringilla dui eget dignissim.
+% focus on CPU parallelisation
+\subsection{Communicator}
+% see Presentation
+\subsection{Tags}
+% see Presentation
+\subsection{Partitioned Operations}
+% see Presentation
+\subsection{Sessions} %no MPI Endpoints, since they didn't make the cut for MPI4.0 and are thus not relevant to current parallelisation
 
-Vivamus non ipsum nunc.
-Donec sit amet erat vitae enim vulputate ultricies nec id mauris.
-Phasellus consectetur sem aliquet malesuada egestas.
-Aliquam viverra libero eros, ac scelerisque elit vestibulum non.
-Integer viverra lacus quis nibh vulputate pharetra.
-Vestibulum viverra nulla et facilisis luctus.
-Sed mollis ultrices nulla.
-Vivamus ut ornare diam.
+\section{Working with accelerators}
+% GPU parallelisation (does this fill an extra chapter though, or would it rather be buried in PartitionedOperations)
+% One might write something to MPIX_Stream?
+
+% I'd love to write about actual performance, but I doubt that I can get a working example on Ara in under a week
+% Also I straight up wouldn't know how to benchmark it properly
+
+% End, roughly 0.5 pages
+\section{Conclusions}
+\section{References}
 

### 2024-07-08 HPC/paper.tex
Added Communicator Sample Text and citations

@@
 
+% TODO: Was kann man in so nem Paper großartig zitieren außer dem MPI Standard?
+
 % Introduction, roughly 1 page
@@
 
-Supercomputing relies on multiple layers of parallelisation, one intranode and one internode.
-Historically this was handled fine with a MPI+MPI (or MPI-Everywhere) programming model.
-In modern supercomputing rising core counts neccessitates a more nuanced programming model, one that makes scaling across cores on a node easier.
-This caused MPI+OpenMP to become the new de facto standard programming model in Supercomputing.
-It simplifies the seperation of internode and intranode tasks greatly, but it also imposes new challenges for developers.
-This paper tries to summarize the current (MPI4.0) methods to overcome these challenges and unlock high performance MPI+OpenMP applications.
+% Historisch MPI+MPI
+% heute MPI+OpenMP (warum)
+% Was sind die schwierigkeiten mit MPI+OpenMP
+% Paper versucht Lösungen zusammenzufassen
+Message Passing Interface is the de facto standard for scaling High Performance Computing (HPC) Workloads across nodes.
+Historically applications relied on the memory wasteful model of MPI+MPI (or MPI-Everywhere) to achieve good scalability across nodes.
+In the last decades however a rapid increase in core counts per CPU neccessitated a switch to MPI+OpenMP to better use newer CPU's ressources.
+While this did work, it brought fourth new challenges in designing software for HPC workloads.
+The main pain point is the much harder to achieve parallel communication in MPI+OpenMP compared to MPI+MPI.
+The newest MPI version (4.0) introduced new ways to express communication parallelism in MPI software to ease this pain point.
+So this paper tries to give an overview over the current ways to express parallel communication to build scalable MPI+OpenMP software.
+This includes a brief theoretical look at each method as well as sample code so developers may apply these methods to their own software.
+
 % this feels like making a buzzword list
@@
 % write why parallel communications is such a large burdon
+The historically preferred method of MPI+MPI made logical parallelism easy to express in software.
+It did however have massive problems with memory management.
+Running a MPI+MPI program on a node with $n$ cores meant having $n$ copies of the program in memory.
+It also leads to some less than desirable effects in the MPI library as each node has to maintain a full list of all MPI nodes in \verb|MPI_COMM_WORLD|.
+%todo: this arguably isn't a problem as long as you don't run on millions of nodes, so include it or nah? especially with how MPI_Sessions can remove the need entirely
+This means for a program of m nodes with n cores each, that each node has to maintain a list of $n \cdot n \cdot m = n^2 \cdot m$ MPI nodes.
+Running an equivalent MPI+OpenMP program with 1 OpenMP program per node addresses these problems in an elegant way.
+%todo: citation or test needed
+First of all the whole program no longer needs to be copied $n$ times on each core, but nearly the spawned parallel loop sections of OpenMP, which uses significantly less memory.
+Secondly the size of \verb|MPI_COMM_WORLD| reduced to $m$ MPI nodes with each node only having to keep 1 such list.
+For modern processors with over $n=100$ cores on a dual socket system this leads to a reduction of over $40000$ in list size on each processor, enabling far higher scaling before \verb|MPI_COMM_WORLD| size is a problem again.
+
+MPI+OpenMP programs do however typically exhibit parallelisation problems in their communication.
+That is caused by the MPI libraries typically being very conservative with their optimizations.
+The MPI standard also mandates a high amount of serialization, for example all requests on the same \verb|<Communicator,Rank>| pair need to be serialized as per the standard.
+That requirement is important for serial MPI programs (as you'd have in MPI+MPI for example) as it prevents deadlocks where two or more nodes wait on each other.
+In multithreaded OpenMP programs it prohibits trivial parallelisation where each Node just communicates over \verb|MPI_COMM_WORLD|.
+So instead developers have to fall back on more sophisticated techniques, which is the topic of the rest of this paper.
 
@@
 
-% focus on CPU parallelisation
+% should there be a explainer for all versions before or inlined?
+
+In contrast to older versions MPI4.0 introduced new operations (Partitioned Operations, Sessions) explicitely for parallel communication \cite{mpi40}.
+%todo: gab es das relaxing zu tags erst in 4.0 oder schon verher? wenn letzteres, dann folgenden Satz entfernen.
+This together with added hints to relax matching requirements makes MPI4.0 more suited towards fast MPI+OpenMP implementations.
+%todo: subsections?; is it all current ways or are there more in the standard? if yes list them or ignore? write for MPI4.0 or MPI4.1? 4.2?
+In the following subsections this paper will discuss the current ways to parallelise communication between nodes in MPI4.0.
+
 \subsection{Communicator}
-% see Presentation
+
+%todo: is simple the right word? most backwards compatible?
+The simplest though maybe not most intuitive way to parallelise communication is to assign each parallel communication channel its own communicator.
+This works since MPI only imposes serialization on the same \verb|<Communicator,Rank>| pair by default.
+That same statement implies no relative ordering between different Communicators \cite{zambreLessonsLearned2022}, hence its potential for expressing parallelisation.
+To use different Communicator the developer needs to be aware of a few key problems:
+
+1) For optimal parallelisation there needs to be two Communicator for each pair of threads communicating internode.
+%todo: insert figure, also link figure
+To demonstrate the need for two Communicator per pair of threads imagine a problem devided into a line of Nodes.
+If only one Communicator is used per pair of threads, communication between \verb|<Rank_0,Thread_5>| and \verb|<Rank_1,Thread_3>| shares its Communicator with the communication between \verb|<Rank_1,Thread_5>| and \verb|<Rank_2,Thread_3>|.
+This is problematic since the messages from \verb|Rank_0| and \verb|Rank_2| to \verb|Rank_1| now run on the same communicator and with that those two exchanges are serialized.
+
+2) For large multidimensional stencils the number of Communicators needed explodes and you should consider alternative parallelisation methods.
+%todo: mark AMD EPYC Rome as Name?
+For more complex stencils like a 27-point stencil as seen in Hypre\cite{hypre2020} on a larger processor like a 64-core AMD EPYC Rome the number of Communicators needed to express parallelism amounts to 808 despite only needing 56 parallel channels for an optimal mapping \cite{zambreLessonsLearned2022}.
+This not only makes it harder to develop the optimal Communicator mapping, it also strains the MPI library that has to map the multitude of Communicators to a far smaller number of hardware contexts (for example 160 in Omni-Path \cite{intelOmniPath}).
+Since the library lacks the knowledge of the intended communication pattern its mapping to the relevant hardware contexts will be suboptimal.
+
+3) Using Communicators for anything but parallelism introduces either performance loss or limits the portability of the program.
+Since MPI4.0 lacks standardised hints regarding Communicator usage, MPI libraries need to assume every Communicator is used for expressing parallelism if it parallelises Communicators at all .
+In such a case, if any Communicator is not used for parallelisation, it puts additional strain on the library for no benefit reducing overall performance.
+%todo: find example hint in MPICH or OpenMP
+A software may be written with library specific hints regarding this usage, but doing so binds one to that library to achieve the desired performance \cite{zambreLessonsLearned2022}.
+
+4) Using Communicators for irregular access patterns degrades performance.
+Imagine a taskbased application with $n$ worker threads and $1$ master thread on each node.
+%todo: generate new graphic like "Lessons Learned on MPI+Threads" Fig. 5 but with N nodes to better demonstrate the infeasability of "Duplicating Communicators per communication partner"
+All workers on one node may need to communicate with the master thread of any other node.
+In such a case the software may either generate $n$ Communicator for each worker thread on a node or generate $n \cdot m$ Communicator for all $n$ worker on a node and all $m$ nodes.
+The latter version is infeasable for any non trivial number of nodes as it strains the MPI library for the same reason large multidimensional stencils do.
+So the software is realistically restricted to just $n$ Communicators.
+This then forces the master thread to communicate over teh same Communicators as its workers.
+And as the master thread needs to iterate over all possible Communicators it will contend with its workers to access the Communicators and slow down the application that way.
+This leads to Communicators performing worse than other parallelisation alternatives.
+As example it performed 1.63x worse than Tags with hints in the task-based framework Legion \cite{zambreLogicalParallel2021}.
+
 \subsection{Tags}
@@
 
-\section{Working with accelerators}
-% GPU parallelisation (does this fill an extra chapter though, or would it rather be buried in PartitionedOperations)
-% One might write something to MPIX_Stream?
+\section{Lorem Ipsum}
 
-% I'd love to write about actual performance, but I doubt that I can get a working example on Ara in under a week
-% Also I straight up wouldn't know how to benchmark it properly
+Anmerkung Justus: Der folgende Lorem Ipsum Text ist nur eingefügt damit ich ein Gefühl fürs Layout kriege, er ist natürlich nicht Teil des finalen Papers.
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Donec et magna ac felis condimentum molestie sed eu nunc.
+Nulla dignissim erat a enim faucibus scelerisque.
+Maecenas sed nisl a orci aliquam tempus sit amet et ligula.
+Quisque vel rhoncus odio.
+Aenean euismod, quam a suscipit placerat, nisl eros viverra arcu, et vehicula arcu arcu ac nibh.
+Proin in erat ac est semper mollis in in neque.
+Sed vitae tellus tincidunt, sodales risus non, aliquam nunc.
+Fusce at faucibus arcu.
+Fusce vulputate mauris eros, at tempor magna pharetra eu.
+Vivamus ut lorem vitae ipsum porta egestas.
+Fusce pretium erat at risus suscipit ultricies.
+Ut mollis, nunc nec tincidunt condimentum, nulla eros pretium magna, ut pulvinar augue turpis sit amet ipsum.
+Sed a eros at eros tristique dictum.
+Pellentesque suscipit libero pulvinar, finibus arcu et, gravida arcu.
+
+Nunc ac dui ullamcorper, consequat velit at, porta eros.
+Ut ac dignissim nulla.
+Phasellus tincidunt consectetur iaculis.
+Aliquam dignissim mauris eu augue faucibus, quis mollis nulla scelerisque.
+Morbi enim diam, tempor et elit sit amet, mollis lacinia orci.
+Phasellus ornare, sapien at dapibus vehicula, turpis libero bibendum sem, vitae lobortis sem elit et erat.
+Vivamus pharetra nunc id nunc molestie dignissim.
+Sed tempor bibendum ipsum, ac volutpat ligula vestibulum non.
+Sed luctus luctus diam vel faucibus.
+Vestibulum in facilisis justo, vitae aliquam nunc.
+Nunc lacinia vehicula arcu.
+Curabitur luctus ex vel quam ultrices gravida.
+Mauris elementum sem sit amet nisi hendrerit consectetur id et elit.
+Donec aliquam eros vitae ligula malesuada egestas.
+
+Maecenas elit tellus, hendrerit sit amet lectus at, suscipit posuere lacus.
+Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
+Quisque vitae vestibulum erat.
+Phasellus aliquet ligula nec molestie sagittis.
+Nam et neque vehicula, accumsan sem vitae, efficitur nulla.
+Maecenas sit amet erat sapien.
+In scelerisque ligula sit amet ante gravida rhoncus.
+Etiam non tempor libero.
+Praesent efficitur imperdiet rutrum.
+Aliquam non dapibus est.
+Suspendisse potenti.
+Sed ac malesuada nisl.
+Mauris semper id ex sit amet mollis.
+Nulla massa justo, volutpat id orci et, auctor dictum magna.
+Suspendisse tincidunt ultrices tortor, vitae mattis ligula tempor non.
+Suspendisse sollicitudin suscipit mi, vel pulvinar est semper in.
+
+Quisque eleifend vehicula lorem eget placerat.
+Nullam malesuada odio non porttitor porttitor.
+Aliquam gravida molestie ligula, at bibendum ipsum sollicitudin sit amet.
+Duis efficitur posuere massa, sit amet ultricies urna elementum eget.
+Donec id mauris vitae sem consectetur finibus et at justo.
+Donec consectetur nisi non nunc viverra, mattis ultricies eros ultricies.
+Phasellus quis blandit tellus.
+Sed sodales consequat erat nec rutrum.
+Suspendisse efficitur nibh neque.
+Praesent eget placerat lectus.
+Nam pretium neque vel lectus aliquam egestas.
+Nulla at placerat quam, vitae tempor metus.
+Etiam rutrum lacus in metus congue, ut tincidunt sapien porttitor.
+Cras id orci ut leo ornare commodo at dapibus metus.
+
+Vivamus bibendum, odio id egestas fermentum, diam libero porta lectus, et finibus turpis tortor quis magna.
+Mauris lorem neque, porta ut dui vel, pellentesque pulvinar nisl.
+Ut nulla nisl, hendrerit sed rutrum ac, sodales in dui.
+Integer scelerisque felis arcu, a iaculis massa varius ut.
+Donec venenatis fringilla magna, sed convallis sem sollicitudin id.
+Fusce mi elit, ultricies nec laoreet eget, condimentum id nibh.
+Maecenas aliquam elit quis lectus consectetur hendrerit.
+Duis ut viverra sem, ac lobortis ex.
+Curabitur viverra volutpat quam finibus ornare.
+
+Pellentesque aliquet tellus non erat fermentum, commodo vestibulum dolor pulvinar.
+Etiam scelerisque tortor at tincidunt vulputate.
+Etiam lobortis nisl erat, ac imperdiet magna semper sit amet.
+Sed at dictum ex.
+Aliquam dictum, nisl elementum rutrum feugiat, nisl sapien rutrum augue, quis lacinia nisi massa a mauris.
+Curabitur fermentum molestie leo vitae faucibus.
+Phasellus sed ligula nibh.
+Nam non pharetra nibh, a interdum ipsum.
+Praesent consequat massa placerat, ornare erat eget, rutrum felis.
+Praesent vel ex sagittis, euismod sem ut, feugiat augue.
+
+Nam ac sapien id justo vulputate tempus eget nec sapien.
+Pellentesque nec enim dignissim, dapibus elit vel, aliquam lectus.
+Etiam ultrices elementum magna at lacinia.
+Phasellus ac lacus a massa ullamcorper imperdiet eget non turpis.
+Aliquam euismod tellus in turpis porttitor sollicitudin.
+Phasellus consectetur, arcu a luctus sodales, tellus ex aliquet odio, in pharetra felis augue et lectus.
+Nam tristique nec elit vitae tristique.
+Nunc orci purus, venenatis lobortis pellentesque auctor, pellentesque vitae tortor.
+Sed metus elit, accumsan at auctor non, convallis vel tortor.
+Ut sagittis iaculis dapibus.
+Fusce est ligula, maximus sed sagittis eu, lobortis at nisl.
+Vestibulum sollicitudin, odio ac cursus fermentum, ante nulla dignissim nisl, ac porttitor felis nisi eget nisl.
+Aenean id purus eu turpis lobortis feugiat.
+In ipsum erat, varius sed orci at, egestas rhoncus nisi.
+Pellentesque vulputate porta urna et tristique.
+In vel dui a nisi rhoncus varius ac vel metus.
+
+Cras tincidunt justo urna.
+Nulla purus erat, elementum a nisi et, sollicitudin vestibulum dui.
+Mauris tristique, augue sed viverra dignissim, velit ligula euismod lectus, sed sollicitudin velit nunc non mi.
+Sed sodales tellus in arcu mollis, eget porttitor turpis pulvinar.
+Nunc efficitur aliquam purus, vel venenatis nisl maximus vitae.
+Donec cursus arcu dolor, et accumsan libero auctor ac.
+Vestibulum viverra massa ac magna vestibulum, sed fringilla ipsum pulvinar.
+Aenean id libero posuere, ultrices elit vel, sagittis risus.
+
+Quisque lorem mauris, semper in nibh sed, imperdiet rutrum justo.
+Sed risus odio, egestas et dolor quis, euismod tempus neque.
+Morbi mattis fermentum nunc, id commodo risus pharetra at.
+Etiam accumsan sapien lorem, eu feugiat arcu blandit at.
+Ut tincidunt id odio ac interdum.
+Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia curae; Mauris euismod sem magna, a pretium risus venenatis non.
+Nunc accumsan orci eu justo tristique tincidunt.
+Suspendisse nibh justo, condimentum eget cursus sed, blandit venenatis quam.
+Phasellus eros nulla, rhoncus eu hendrerit eu, vestibulum in ligula.
+Vestibulum nec mollis neque.
+Aliquam vel mauris et nulla pulvinar sodales.
+Sed tempor sapien turpis, vitae viverra ligula viverra vitae.
+Etiam efficitur congue lectus, ac semper velit semper a.
+
+In viverra mollis aliquet.
+Suspendisse finibus justo quis orci lacinia, sed pharetra sem congue.
+Mauris at erat lectus.
+Duis ipsum massa, vulputate non felis rutrum, aliquet aliquet augue.
+Donec molestie mi a ipsum accumsan faucibus.
+Etiam lacinia nunc sed varius congue.
+Phasellus et sollicitudin diam, sit amet vulputate leo.
+Pellentesque non lorem sodales nisi luctus sollicitudin.
+
+Cras posuere nunc at nulla auctor fermentum.
+Quisque ullamcorper dolor neque, ac rutrum magna dapibus et.
+Morbi ut nisl nec leo hendrerit malesuada.
+Vivamus orci nisl, molestie ac sem ac, elementum viverra augue.
+Cras quis venenatis lacus.
+Vestibulum quis rhoncus urna.
+Suspendisse libero neque, euismod vel molestie eget, lacinia ac enim.
+Nulla ultrices tempus bibendum.
+Donec eu felis mauris.
+Sed eget ligula et purus lacinia fringilla eget quis ligula.
+
+Morbi venenatis consectetur erat, finibus finibus leo.
+Proin eleifend nec magna quis iaculis.
+Vestibulum sit amet odio gravida, ornare ante aliquet, mattis dui.
+Ut lobortis mauris sed semper aliquet.
+Suspendisse fermentum pretium ex eu pellentesque.
+Aenean ac auctor enim, quis egestas nisi.
+Duis at tincidunt felis.
+Mauris et dolor sit amet lectus auctor sollicitudin vel ultricies nibh.
+Morbi interdum nunc quam, in pellentesque nunc aliquam quis.
+Cras blandit orci ac purus posuere mollis.
+Duis scelerisque blandit mi sed blandit.
+Phasellus mollis interdum ante, vitae congue leo dignissim eget.
+Nam ac orci eget elit rutrum sodales id sit amet mi.
+
+Nulla scelerisque condimentum commodo.
+Vivamus scelerisque faucibus metus non viverra.
+Curabitur eget mauris feugiat, tempor tortor sit amet, blandit leo.
+Duis nec enim sagittis, tempor tortor auctor, vestibulum arcu.
+Nunc pretium, sem sed vestibulum tristique, orci tellus mattis nulla, in dictum sem tellus in orci.
+Morbi at massa diam.
+Vivamus vehicula, ligula ac tristique eleifend, elit enim porta urna, ac efficitur massa turpis a purus.
+Sed tempus massa ac orci mattis, sit amet convallis turpis porttitor.
+Maecenas scelerisque velit vel orci feugiat, ac viverra erat dictum.
+Nullam quis lectus in dolor eleifend ultricies vel eu magna.
+
+Donec varius aliquam cursus.
+Maecenas mattis nisl sit amet nibh sollicitudin finibus.
+Pellentesque porta sed orci nec elementum.
+Nam ac urna ultricies felis hendrerit finibus.
+Duis velit mi, ullamcorper a pellentesque sit amet, interdum nec velit.
+Vivamus eu libero erat.
+Aliquam erat volutpat.
+Nam dignissim enim nibh, vel mattis nibh tempus vitae.
+Aliquam congue nulla eget sapien commodo, id auctor dolor tristique.
+Aenean vel sem fringilla dui hendrerit pharetra sed vel elit.
+
+%vll. Wort zur Umsetzung in Bibliotheken? OpenMPI und MPICH haben es beide immerhin ...
 
@@
 \section{Conclusions}
-\section{References}
+\bibliographystyle{plain} % We choose the "plain" reference style
+\bibliography{citations} % Entries are in the refs.bib file
 

### 2024-07-12 HPC/paper.tex
Added ToDo
Added ACM Reference Format
Updated Motivation

@@
-\documentclass[sigconf]{acmart} %\documentclass[sigconf, language=german]{acmart}
+\documentclass[sigconf]{acmart}
 \usepackage{booktabs}
@@
 
-\title{MPI+OpenMP in MPI4.0}
+%todo: look into title
+\title{Parallelising Hybrid MPI Programms efficiently}
 \author{Dreßler Justus}
@@
 
-% TODO: Was kann man in so nem Paper großartig zitieren außer dem MPI Standard?
-
 % Introduction, roughly 1 page
+% todo: rework
 \section{Introduction}
 
-% Historisch MPI+MPI
+% Historisch MPI-Everywhere
 % heute MPI+OpenMP (warum)
@@
 Message Passing Interface is the de facto standard for scaling High Performance Computing (HPC) Workloads across nodes.
-Historically applications relied on the memory wasteful model of MPI+MPI (or MPI-Everywhere) to achieve good scalability across nodes.
+Historically applications relied on the memory wasteful model of MPI-Everywhere to achieve good scalability across nodes.
 In the last decades however a rapid increase in core counts per CPU neccessitated a switch to MPI+OpenMP to better use newer CPU's ressources.
 While this did work, it brought fourth new challenges in designing software for HPC workloads.
-The main pain point is the much harder to achieve parallel communication in MPI+OpenMP compared to MPI+MPI.
+The main pain point is the much harder to achieve parallel communication in MPI+OpenMP compared to MPI-Everywhere.
 The newest MPI version (4.0) introduced new ways to express communication parallelism in MPI software to ease this pain point.
@@
 \section{Motivation}
-% write why parallel communications is such a large burdon
-The historically preferred method of MPI+MPI made logical parallelism easy to express in software.
+The historically preferred method of MPI-Everywhere made logical parallelism easy to express in software.
+%todo: cite or rework
 It did however have massive problems with memory management.
-Running a MPI+MPI program on a node with $n$ cores meant having $n$ copies of the program in memory.
-It also leads to some less than desirable effects in the MPI library as each node has to maintain a full list of all MPI nodes in \verb|MPI_COMM_WORLD|.
-%todo: this arguably isn't a problem as long as you don't run on millions of nodes, so include it or nah? especially with how MPI_Sessions can remove the need entirely
-This means for a program of m nodes with n cores each, that each node has to maintain a list of $n \cdot n \cdot m = n^2 \cdot m$ MPI nodes.
-Running an equivalent MPI+OpenMP program with 1 OpenMP program per node addresses these problems in an elegant way.
-%todo: citation or test needed
-First of all the whole program no longer needs to be copied $n$ times on each core, but nearly the spawned parallel loop sections of OpenMP, which uses significantly less memory.
-Secondly the size of \verb|MPI_COMM_WORLD| reduced to $m$ MPI nodes with each node only having to keep 1 such list.
-For modern processors with over $n=100$ cores on a dual socket system this leads to a reduction of over $40000$ in list size on each processor, enabling far higher scaling before \verb|MPI_COMM_WORLD| size is a problem again.
+Classical MPI-Everywhere programs spawn a single process per core available.
+MPI+OpenMP programs normally spawn a single process per node or NUMA domain \cite{zambreLessonsLearned2022}.
 
-MPI+OpenMP programs do however typically exhibit parallelisation problems in their communication.
-That is caused by the MPI libraries typically being very conservative with their optimizations.
-The MPI standard also mandates a high amount of serialization, for example all requests on the same \verb|<Communicator,Rank>| pair need to be serialized as per the standard.
-That requirement is important for serial MPI programs (as you'd have in MPI+MPI for example) as it prevents deadlocks where two or more nodes wait on each other.
-In multithreaded OpenMP programs it prohibits trivial parallelisation where each Node just communicates over \verb|MPI_COMM_WORLD|.
-So instead developers have to fall back on more sophisticated techniques, which is the topic of the rest of this paper.
+%todo: cite MPI 3 Standard
+A MPI communicator is a construct that defines which processes may talk with each other.
+Each MPI process is required to keep a list of all participating MPI processes in the default communicator \verb|MPI_COMM_WORLD|.
+Assuming an $n$ core processor $n$ copies of the list are lying in each nodes memory in the MPI-Everywhere model.
+An equivalent MPI+OpenMP program shares its list between its cores and also spawns $n$ times less processes, so \verb|MPI_COMM_WORLD| is $n$ times smaller and $n$ times less often on each processor.
+For current generation processors exceeding 100 cores like Intel® Xeon® 6 \cite{intelXeon6} or 4th Generation AMD EPYC™ \cite{amd4thGenEpyc} this equates to a reduction of \verb|MPI_COMM_WORLD|'s memory consumption of over $n^2 > 10.000$ times.
+
+Each MPI process also needs to be a full program, which means each core has to have all instructions duplicated.
+OpenMP however only needs to generate code for the actually parallised parts (typically loops) of the program, which reduces the size for all additional cores significantly.
+
+MPI+OpenMP programs do, however, tend to perform slower than MPI-Everywhere versions in practice \cite{zambreLessonsLearned2022}.
+This is caused in part by accidental synchronization while calling the MPI library.
+%todo: cite standard
+The MPI standard mandates a high amount of serialization to ensure correct communication, for example all messages on the same \verb|<Communicator,Rank>| pair are defined as non-overtaking in MPI 4.0 Section 3.5\cite{mpi40}.
+This is important to prevent deadlocks (two processes each wait on the other) but also slows down the communication overall.
+We will explore how to circumvent such slowdowns in the following section.
 
 % Body, roughly 2.5 pages
-\section{Parallel Communication in MPI4.0}
+\section{Parallel Communication in MPI 4.0}
 
@@
 
-In contrast to older versions MPI4.0 introduced new operations (Partitioned Operations, Sessions) explicitely for parallel communication \cite{mpi40}.
+In contrast to older versions MPI 4.0 introduced new operations (Partitioned Operations, Sessions) explicitely for parallel communication \cite{mpi40}.
 %todo: gab es das relaxing zu tags erst in 4.0 oder schon verher? wenn letzteres, dann folgenden Satz entfernen.
-This together with added hints to relax matching requirements makes MPI4.0 more suited towards fast MPI+OpenMP implementations.
-%todo: subsections?; is it all current ways or are there more in the standard? if yes list them or ignore? write for MPI4.0 or MPI4.1? 4.2?
-In the following subsections this paper will discuss the current ways to parallelise communication between nodes in MPI4.0.
+This together with added hints to relax matching requirements makes MPI 4.0 more suited towards fast MPI+OpenMP implementations.
+In the following subsections this paper will discuss the current ways to parallelise communication between nodes in MPI 4.0.
 
@@
 %todo: insert figure, also link figure
+%todo: possibly code for that same figure so the complexity can be better judged by a reader?
 To demonstrate the need for two Communicator per pair of threads imagine a problem devided into a line of Nodes.
@@
 3) Using Communicators for anything but parallelism introduces either performance loss or limits the portability of the program.
-Since MPI4.0 lacks standardised hints regarding Communicator usage, MPI libraries need to assume every Communicator is used for expressing parallelism if it parallelises Communicators at all .
+Since MPI 4.0 lacks standardised hints regarding Communicator usage, MPI libraries need to assume every Communicator is used for expressing parallelism if it parallelises Communicators at all .
 In such a case, if any Communicator is not used for parallelisation, it puts additional strain on the library for no benefit reducing overall performance.
@@
 % see Presentation
-\subsection{Sessions} %no MPI Endpoints, since they didn't make the cut for MPI4.0 and are thus not relevant to current parallelisation
+\subsection{Sessions}
 
@@
 
-Morbi venenatis consectetur erat, finibus finibus leo.
-Proin eleifend nec magna quis iaculis.
-Vestibulum sit amet odio gravida, ornare ante aliquet, mattis dui.
-Ut lobortis mauris sed semper aliquet.
-Suspendisse fermentum pretium ex eu pellentesque.
-Aenean ac auctor enim, quis egestas nisi.
-Duis at tincidunt felis.
-Mauris et dolor sit amet lectus auctor sollicitudin vel ultricies nibh.
-Morbi interdum nunc quam, in pellentesque nunc aliquam quis.
-Cras blandit orci ac purus posuere mollis.
-Duis scelerisque blandit mi sed blandit.
-Phasellus mollis interdum ante, vitae congue leo dignissim eget.
-Nam ac orci eget elit rutrum sodales id sit amet mi.
-
-Nulla scelerisque condimentum commodo.
-Vivamus scelerisque faucibus metus non viverra.
-Curabitur eget mauris feugiat, tempor tortor sit amet, blandit leo.
-Duis nec enim sagittis, tempor tortor auctor, vestibulum arcu.
-Nunc pretium, sem sed vestibulum tristique, orci tellus mattis nulla, in dictum sem tellus in orci.
-Morbi at massa diam.
-Vivamus vehicula, ligula ac tristique eleifend, elit enim porta urna, ac efficitur massa turpis a purus.
-Sed tempus massa ac orci mattis, sit amet convallis turpis porttitor.
-Maecenas scelerisque velit vel orci feugiat, ac viverra erat dictum.
-Nullam quis lectus in dolor eleifend ultricies vel eu magna.
-
-Donec varius aliquam cursus.
-Maecenas mattis nisl sit amet nibh sollicitudin finibus.
-Pellentesque porta sed orci nec elementum.
-Nam ac urna ultricies felis hendrerit finibus.
-Duis velit mi, ullamcorper a pellentesque sit amet, interdum nec velit.
-Vivamus eu libero erat.
-Aliquam erat volutpat.
-Nam dignissim enim nibh, vel mattis nibh tempus vitae.
-Aliquam congue nulla eget sapien commodo, id auctor dolor tristique.
-Aenean vel sem fringilla dui hendrerit pharetra sed vel elit.
-
 %vll. Wort zur Umsetzung in Bibliotheken? OpenMPI und MPICH haben es beide immerhin ...
@@
 \section{Conclusions}
-\bibliographystyle{plain} % We choose the "plain" reference style
-\bibliography{citations} % Entries are in the refs.bib file
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{citations}
 

### 2024-07-12 other/noten.py

@@ -19,6 +19,7 @@ noten = [
     (1.3,6), # Numerische Mathematik
     # C
     (1.3,6), # IT-Sicherheit 
+    (2.7,6), # Algorithmisches Beweisen
     (1.3,3), # Projektmanagement (ASQ) 
     # A
     (1.0,9), # Grundlagen informatischer Problemlösung 

### 2024-07-12 HPC/paper.tex
Fixed Communicators

--- 2024-07-07 HPC/paper.tex
@@
 % Introduction, roughly 1 page
+% todo: rework
 \section{Introduction}
 
-Supercomputing relies on multiple layers of parallelisation, one intranode and one internode.
-Historically this was handled fine with a MPI+MPI (or MPI-Everywhere) programming model.
-In modern supercomputing rising core counts neccessitates a more nuanced programming model, one that makes scaling across cores on a node easier.
-This caused MPI+OpenMP to become the new de facto standard programming model in Supercomputing.
-It simplifies the seperation of internode and intranode tasks greatly, but it also imposes new challenges for developers.
-This paper tries to summarize the current (MPI4.0) methods to overcome these challenges and unlock high performance MPI+OpenMP applications.
+% Historisch MPI-Everywhere
+% heute MPI+OpenMP (warum)
+% Was sind die schwierigkeiten mit MPI+OpenMP
+% Paper versucht Lösungen zusammenzufassen
+Message Passing Interface is the de facto standard for scaling High Performance Computing (HPC) Workloads across nodes.
+Historically applications relied on the memory wasteful model of MPI-Everywhere to achieve good scalability across nodes.
+In the last decades however a rapid increase in core counts per CPU neccessitated a switch to MPI+OpenMP to better use newer CPU's ressources.
+While this did work, it brought fourth new challenges in designing software for HPC workloads.
+The main pain point is the much harder to achieve parallel communication in MPI+OpenMP compared to MPI-Everywhere.
+The newest MPI version (4.0) introduced new ways to expose communication parallelism in MPI software to ease this pain point.
+So this paper tries to give an overview over the current ways to expose parallel communication to build scalable MPI+OpenMP software.
+This includes a brief theoretical look at each method as well as sample code so developers may apply these methods to their own software.
+
 % this feels like making a buzzword list
 
+% Problemstellung: Wie expose ich logisch Parallele Nachrichten an eine MPI Biblithek (MPICH, OpenMPI) sodass diese es auf parallele Kommunikationswege mappen kann?
+
 \section{Motivation}
-% write why parallel communications is such a large burdon
+The historically preferred method of MPI-Everywhere made logical parallelism easy to express in software.
+%todo: cite or rework
+It did however have massive problems with memory management.
+Classical MPI-Everywhere programs spawn a single process per core available.
+MPI+OpenMP programs normally spawn a single process per node or NUMA domain \cite{zambreLessonsLearned2022}.
+
+%todo: cite MPI 3 Standard
+A MPI communicator is a construct that defines which processes may talk with each other.
+Each MPI process is required to keep a list of all participating MPI processes in the default communicator \verb|MPI_COMM_WORLD|.
+Assuming an $n$ core processor $n$ copies of the list are lying in each nodes memory in the MPI-Everywhere model.
+An equivalent MPI+OpenMP program shares its list between its cores and also spawns $n$ times less processes, so \verb|MPI_COMM_WORLD| is $n$ times smaller and $n$ times less often on each processor.
+For current generation processors exceeding 100 cores like Intel® Xeon® 6 \cite{intelXeon6} or 4th Generation AMD EPYC™ \cite{amd4thGenEpyc} this equates to a reduction of \verb|MPI_COMM_WORLD|'s memory consumption of over $n^2 > 10.000$ times.
+
+Each MPI process also needs to be a full program, which means each core has to have all instructions duplicated.
+OpenMP however only needs to generate code for the actually parallised parts (typically loops) of the program, which reduces the size for all additional cores significantly.
+
+MPI+OpenMP programs do, however, tend to perform slower than MPI-Everywhere versions in practice \cite{zambreLessonsLearned2022}.
+This is caused in part by accidental synchronization while calling the MPI library.
+The MPI standard mandates a high amount of serialization to ensure correct communication, for example all messages on the same \verb|<Communicator,Rank>| pair are transmitted in non-overtaking order \cite{mpi40}.
+This is important to prevent deadlocks (two processes each wait on the other) but also slows down the communication overall.
+We will explore how to circumvent such slowdowns in the following section.
 
 % Body, roughly 2.5 pages
-\section{Parallel Communication in MPI4.0}
+\section{Parallel Communication in MPI 4.0}
 
-% focus on CPU parallelisation
+There are several ways to implement logical parallel communication efficiently in MPI.
+For MPI versions lower than 4.0 unique communicators may be used per logical parallel communication channel.
+MPI 4.0 introduced Partitioned Operations to aid efficient parallel communication.
+In Partitioned Operations multiple threads each contribute part of a larger message.
+% todo: check that MPI_ANY_TAG is first documented in MPI 4.0 and not already earlier
+MPI 4.0 also added new hints, a way for a program to guarantee MPI that it will not use certain features, that can make tags viable for performant parallel communication.
+% todo: MPI Sessions yes or no? -> Read up if time, don't forget graphics and code before though
+Furthermore MPI 4.0 introduced MPI Sessions, which may be used for parallelisation purposes, but are not considered in the scope of this paper.
+In the following subsections this paper will discuss the mentioned methods in more detail.
+
 \subsection{Communicator}
-% see Presentation
+
+%todo: is simple the right word? most backwards compatible?
+The simplest though maybe not most intuitive way to parallelise communication is to assign each logical parallel message its own communicator.
+Since MPI gives no implies no relative ordering on different communicators \cite{zambreLessonsLearned2022}, libraries can map messages on different communicators to different underlying message channels.
+To use different communicator an MPI application developer needs to be aware of a few key problems:
+
+1) For optimal parallelisation there needs to be two communicator for each pair of threads communicating between different processes.
+%todo: insert figure, also link figure
+%todo: possibly code for that same figure so the complexity can be better judged by a reader?
+To demonstrate the need for two communicator per pair of threads imagine a problem devided into a line of Nodes.
+If only one communicator is used per pair of threads, communication between \verb|<Rank_0,Thread_5>| and \verb|<Rank_1,Thread_3>| shares its communicator with the communication between \verb|<Rank_1,Thread_5>| and \verb|<Rank_2,Thread_3>|.
+This is problematic since the messages from \verb|Rank_0| and \verb|Rank_2| to \verb|Rank_1| now run on the same communicator and with that those two exchanges are no longer independent.
+
+2) For large multidimensional stencils the number of communicators needed grows rapidly and an MPI program using those should consider other methods to expose parallelisation.
+For more complex stencils like a 27-point 3D stencil as seen in Hypre\cite{hypre2020} on a larger 64-core processor (data distributed as [4,4,4] chunks) the number of communicators needed to expose all possible communication parallelism amounts to $6 \cdot 4^2$ for sides plus $24 \cdot 4^2 - 8$ for the corner diagonal plus $24 \cdot 4^2 - 48$ for the edge diagonals, totalling $808$ communicators.
+The workload would only need a single parallel communication channel for each thread communicating between nodes ideally though, which is only $4^3 - (4-2)^3 = 56$.
+The equations were derived from those used by Zambre and Chandramowlishwaran \cite{zambreLessonsLearned2022}.
+Generating so many communicators is not only complex, it also strains MPI libraries that have to map the many possibly parallel messages to a far smaller number of parallel paths in hardware (as an example Omni-Path only has 160 hardware contexts \cite{intelOmniPath}).
+Since the MPI library lacks the knowledge of the intended communication pattern its mapping to the relevant hardware contexts will be suboptimal.
+
+3) Using communicators to expose parallelism and group processes introduces performance loss or limits the portability across MPI libraries.
+Since MPI 4.0 lacks standardised hints regarding communicator usage, MPI libraries cannot differentiate between communicators used to expose parallelism and those that are used for grouping processes together.
+If a MPI library tries to parallelise across communicators, any communicator that is not used to expose parallelisation puts unneccessary strain on the library, reducing overall performance.
+%todo: find example hint in MPICH or OpenMP
+A program may use library specific hints to specify which communicators are used for what purpose, but doing so makes the program dependent on that MPI library instead of being compatible with any MPI library, at least for the optimal performance \cite{zambreLessonsLearned2022}.
+
+4) Using communicators for irregular access patterns degrades performance.
+Imagine a taskbased application with $n$ worker threads and one main thread in each process.
+%todo: generate new graphic like "Lessons Learned on MPI+Threads" Fig. 5 but with N nodes to better demonstrate the infeasability of "Duplicating communicators per communication partner"
+All workers in one process may need to communicate with the main thread of any other process.
+In such a case the software may either generate $n$ communicator for each worker thread in a process or generate $n \cdot m$ communicator for all $n$ worker in a process and all $m$ processes.
+The latter version is infeasable for any non trivial number of processes as it strains the MPI library for the same reason large multidimensional stencils do.
+So the software is realistically restricted to just $n$ communicators.
+This then forces the main thread to communicate over the same communicators as its workers.
+And as the main thread needs to iterate over all possible communicators it will contend with its workers to access the communicators and slow down the application that way.
+This leads to communicators performing worse than other parallelisation alternatives.
+As example communicators performed 1.63x worse than tags with hints in the task-based framework Legion \cite{zambreLogicalParallel2021}.
+
 \subsection{Tags}
@@
 % see Presentation
-\subsection{Sessions} %no MPI Endpoints, since they didn't make the cut for MPI4.0 and are thus not relevant to current parallelisation
+% \subsection{Sessions}
 
-\section{Working with accelerators}
-% GPU parallelisation (does this fill an extra chapter though, or would it rather be buried in PartitionedOperations)
-% One might write something to MPIX_Stream?
-
-% I'd love to write about actual performance, but I doubt that I can get a working example on Ara in under a week
-% Also I straight up wouldn't know how to benchmark it properly
+%vll. Wort zur Umsetzung in Bibliotheken? OpenMPI und MPICH haben es beide immerhin ...
 
@@
 \section{Conclusions}
-\section{References}
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{citations}
 

### 2024-07-13 HPC/paper.tex
Added Tags with Hints

@@
 
-\subsection{Tags}
+\subsection{Tags with Hints}
+
+MPI operations like \verb|MPI_Send| take a \verb|<Rank,Tag,Communicator>| triple as input.
+The tag is sent to differentiate messages between the same processes on a single communicator.
+Normally the MPI standard requires a non-overtaking order for all messages with the same \verb|<Rank,Communicator>| \cite{mpi40}.
+So the default settings do not enable a program to use the additionally sent tags for parallelisation.
+If the program however guarantees MPI, that it will not use the \verb|MPI_ANY_TAG| (a parameter matching all tags), MPI libraries can send different messages between the same ranks but differing tags in arbitrary order.
+
+To use this new avenue to expose parallelism, programs can simply encode the source thread and target thread into the tag.
+A possible implementation is given in FIG. %todo: add figure of sample code
+Using tags comes with some caveats:
+
+As with using communicators to expose parallelism, tags are also used in other contexts than parallelisation.
+So as with communicators a program has to rely on MPI library specific extra hints to clarify which parts of the tag encodes the parallelisation information or risk a suboptimal distribution to the hardware communication channels.
+An example of such extra hints is shown in FIG for the MPI library MPICH. %todo: add MPICH figure
+
+Tags are restricted to 64 bits.
+Depending on the programs other uses for tags this may be a non issue or a dealbreaker.
+Encoding source and target thread in tags for $n$ core processors takes $2\log_2{n}$ bits.
+For a 128 core processors this would mean, that a program reserves $2\log_2{128} = 14$ bits of their tags.
+For most programs this will be irrelevant, but if the program already encodes much information in their tags, exposing parallelisation information might not be feasable.
+
 % see Presentation

### 2024-07-14 HPC/paper.tex
Added Partitioned Operations

@@
 % Body, roughly 2.5 pages
-\section{Parallel Communication in MPI 4.0}
+\section{Parallel Point-to-Point Communication in MPI 4.0}
 
-There are several ways to implement logical parallel communication efficiently in MPI.
-For MPI versions lower than 4.0 unique communicators may be used per logical parallel communication channel.
+Point-to-Point Communication refers to any communication between 2 different processes.
+There are several ways to implement logical parallel point-to-point communication efficiently in MPI.
+For MPI versions lower than 4.0 a unique communicators per logical parallel communication channel may be used.
 MPI 4.0 introduced Partitioned Operations to aid efficient parallel communication.
 In Partitioned Operations multiple threads each contribute part of a larger message.
-% todo: check that MPI_ANY_TAG is first documented in MPI 4.0 and not already earlier
 MPI 4.0 also added new hints, a way for a program to guarantee MPI that it will not use certain features, that can make tags viable for performant parallel communication.
@@
 
-To use this new avenue to expose parallelism, programs can simply encode the source thread and target thread into the tag.
+To use this new avenue to expose parallelism, programs need to use the \verb|mpi_assert_no_any_tag| hint.
+Now any MPI library is free to parallelise over different tags.
+To use that a program can simply encode the source thread and target thread into the tag.
 A possible implementation is given in FIG. %todo: add figure of sample code
@@
 
-% see Presentation
-\subsection{Partitioned Operations}
+\subsection{Partitioned Point-to-Point Communication}
+
+%todo: add sample code
+
+Partitioned Operations are point-to-point operations where the message is partioned into multiple parts.
+
+To implement partitioned point-to-point operations is slightly more difficult than simple \verb|MPI_Send| and \verb|MPI_Recv| calls.
+In contrast to those all operations have to be first initialised on sender and receiver side with \verb|MPI_PSend_Init| and \verb|MPI_PRecv_Init| respectively.
+The initialisation may be done long before the operation is executed.
+To start the execution there must be a single call to \verb|MPI_Start|.
+Then each contributing sending thread may write its part of the message in the shared buffer and then mark its part as ready by calling \verb|MPI_Pready|.
+On the receiving side threads may poll for completion of individual partitions of the message by calling \verb|MPI_Parrived|.
+As soon as those return a truthy flag a thread may work on that partitions data.
+There must be a single final call to \verb|MPI_Wait| or \verb|MPI_Test| to complete the operation, before it may be restarted with another call to \verb|MPI_Start|.
+
+This a bit more complicated setup allows an MPI library to transmit the partitions in parallel instead of waiting for all threads to finish working.
+Partitioned operations are ideally suited to very regular communication, as found in stencil applications.
+They are very ill suited for irregular communication though, as the operations have to be initialised before any communication may happen \cite{zambreLessonsLearned2022}.
+Irregular communication is further complicated by the missing ability to provide wildcards (\verb|MPI_ANY_TAG|, \verb|MPI_ANY_SOURCE|) to \verb|MPI_PRecv_Init| \cite{mpi40}, on which modern task-based runtimes like Legion rely \cite{zambreLessonsLearned2022}.
+
 % see Presentation

### 2024-07-14 HPC/paper.tex
Added RMA Operations

@@
 
+%vll. Wort zur Umsetzung in Bibliotheken? OpenMPI und MPICH haben es beide immerhin ...
+
 % Historisch MPI-Everywhere
@@
 % Problemstellung: Wie expose ich logisch Parallele Nachrichten an eine MPI Biblithek (MPICH, OpenMPI) sodass diese es auf parallele Kommunikationswege mappen kann?
+% behandelt nur CPU, keine Acceleratoren
+% reminder: Einleiten was MPI ist, was MPI Bibliotheken sind, populäre nennen (MPICH, OpenMP, IntelMPI?), sagen was OpenMP ist und wofür es genutzt wird, Warum man es zusammen tun sollte,
 
@@
 MPI 4.0 also added new hints, a way for a program to guarantee MPI that it will not use certain features, that can make tags viable for performant parallel communication.
-% todo: MPI Sessions yes or no? -> Read up if time, don't forget graphics and code before though
 Furthermore MPI 4.0 introduced MPI Sessions, which may be used for parallelisation purposes, but are not considered in the scope of this paper.
@@
 
-This a bit more complicated setup allows an MPI library to transmit the partitions in parallel instead of waiting for all threads to finish working.
+This more complicated setup allows an MPI library to transmit the partitions in parallel instead of waiting for all threads to finish working.
 Partitioned operations are ideally suited to very regular communication, as found in stencil applications.
@@
 
-% see Presentation
-% \subsection{Sessions}
+\section{Parallel Remote Memory Access in MPI 4.0}
 
-%vll. Wort zur Umsetzung in Bibliotheken? OpenMPI und MPICH haben es beide immerhin ...
+MPI implements Remote Memory Access (RMA) with memory windows.
+In those windows other processes may put data with \verb|MPI_Put| or get data from with \verb|MPI_Get| or their non-blocking alternatives \cite{mpi40}.
+In contrast to point-to-point communication MPI imposes no order requirement for \verb|MPI_Put| and \verb|MPI_Get| operations \cite{mpi40}.
+So for these operations MPI libraries are free to parallelise them as needed.
+MPI windows also allow atomic \verb|MPI_Accumulate| operations.
+This operation updates a target value by adding a number to it atomically.
+MPI imposes strict ordering on \verb|MPI_Accumulate| operations per default.
+If such a ordering guarantee is not needed, Zambre and Chandramowlishwaran list NWChem as an example \cite{zambreLessonsLearned2022}, it may be relaxed on window creation by specifying another \verb|accumulate_ordering|.
+The MPI standard supports specifying no ordering or only a subset of read-after-read, read-after-write, write-after-read, write-after-write in such a case \cite{mpi40}.
+%todo: add example with no ordering
 

### 2024-07-14 HPC/paper.tex
Added introduction

@@
 %todo: look into title
-\title{Parallelising Hybrid MPI Programms efficiently}
+\title{Efficient Hybrid MPI+OpenMP Programming}
 \author{Dreßler Justus}
@@
 % Paper versucht Lösungen zusammenzufassen
-Message Passing Interface is the de facto standard for scaling High Performance Computing (HPC) Workloads across nodes.
-Historically applications relied on the memory wasteful model of MPI-Everywhere to achieve good scalability across nodes.
-In the last decades however a rapid increase in core counts per CPU neccessitated a switch to MPI+OpenMP to better use newer CPU's ressources.
-While this did work, it brought fourth new challenges in designing software for HPC workloads.
-The main pain point is the much harder to achieve parallel communication in MPI+OpenMP compared to MPI-Everywhere.
-The newest MPI version (4.0) introduced new ways to expose communication parallelism in MPI software to ease this pain point.
-So this paper tries to give an overview over the current ways to expose parallel communication to build scalable MPI+OpenMP software.
-This includes a brief theoretical look at each method as well as sample code so developers may apply these methods to their own software.
+Message Passing Interface is the de facto standard for scaling High Performance Computing (HPC) Workloads across multiple compute nodes.
+This paper will speak of MPI as the standard and refer to MPI implementations like MPICH and OpenMPI as "MPI libraries" going forward.
+These MPI libaries implement the MPI standard while trying to achieve maximum performance while adhearing to it.
+To do so some libraries include extra library specific settings as hints, which the library user can set to increase possible performance.
+Sometimes these libraries will also include entirely new ways to parallelise communication like MPIX\_Stream \cite{Zhou2022}, but this paper will focus mainly on the MPI standard itself.
 
-% this feels like making a buzzword list
+Traditionally applications relied on the memory-hungry MPI-Everywhere model \cite{zambreLessonsLearned2022}.
+In the last decades however a rapid increase in core counts per CPU relative to other on-node resources neccessitated a switch to a model of MPI+X,
+where MPI is used to parallise between different nodes and X is used to parallelise on each node.
+Some popular choices for this second X part are CPU multithreading libraries like OpenMP or POSIX Threads, GPU Acceleration with CUDA or using MPI with Remote Memory Access calls.
+This paper will focus on MPI+OpenMP, though many aspects of this paper also apply for any other CPU multithreading library.
 
-% Problemstellung: Wie expose ich logisch Parallele Nachrichten an eine MPI Biblithek (MPICH, OpenMPI) sodass diese es auf parallele Kommunikationswege mappen kann?
-% behandelt nur CPU, keine Acceleratoren
-% reminder: Einleiten was MPI ist, was MPI Bibliotheken sind, populäre nennen (MPICH, OpenMP, IntelMPI?), sagen was OpenMP ist und wofür es genutzt wird, Warum man es zusammen tun sollte,
+OpenMP is a popular library for generating shared-memory multithreaded programs.
+It works under the fork join model, where a main thread does all the serial work and additional threads get spawned for multithreading when needed.
+It is easy to add to existing serial code, mostly by adding comments called pragmas in front of loops.
+Its ease of use also makes it popular for developing HPC software.
+MPI+OpenMP programs tend to perform worse than their MPI-Everywhere counterparts \cite{zambreLessonsLearned2022}\cite{zambreLogicalParallel2021}.
+The main problem being extra unneeded thread synchronization overheads \cite{zambreLessonsLearned2022} and less exposure of parallelism for MPI libraries to use.
+
+This paper gives an overview of ways to expose parallelism in communication to MPI libaries to achieve maximum performance.
+For this it will discuss mainly point-to-point communication.
+It compares traditional methods like generating multiple MPI communicators with the newer features of MPI 4.0, which introduced new hints to make tags a viable method of exposing parallelisation and added new partitioned operations which can gather contributions of multiple threads in parallel.
 
@@
 The historically preferred method of MPI-Everywhere made logical parallelism easy to express in software.
-%todo: cite or rework
-It did however have massive problems with memory management.
+It does however use up more memory and other on chip ressources relative to MPI+OpenMP.
 Classical MPI-Everywhere programs spawn a single process per core available.
@@
 
-%todo: cite MPI 3 Standard
-A MPI communicator is a construct that defines which processes may talk with each other.
-Each MPI process is required to keep a list of all participating MPI processes in the default communicator \verb|MPI_COMM_WORLD|.
+MPI point-to-point communication runs on so called communicators.
+A MPI communicator defines which processes may talk with each other.
+Each MPI process is required to keep a list of all participating MPI processes in the default communicator \verb|MPI_COMM_WORLD| \cite{mpi40}.
 Assuming an $n$ core processor $n$ copies of the list are lying in each nodes memory in the MPI-Everywhere model.
@@
 MPI operations like \verb|MPI_Send| take a \verb|<Rank,Tag,Communicator>| triple as input.
-The tag is sent to differentiate messages between the same processes on a single communicator.
+The tag is included to differentiate messages between the same processes on a single communicator.
 Normally the MPI standard requires a non-overtaking order for all messages with the same \verb|<Rank,Communicator>| \cite{mpi40}.

### 2024-07-15 HPC/paper.tex
Changed Title yet again
Added figures for Communicators and Tags

@@
 \usepackage{booktabs}
+\usepackage{listings}
 
@@
 
-%todo: look into title
-\title{Efficient Hybrid MPI+OpenMP Programming}
+\title{Efficient MPI+OpenMP Programming with MPI 4.0}
 \author{Dreßler Justus}
@@
 
+% todo: revisit?
 \begin{abstract}
 
-    %TODO: add abstract
-    Lorem Ipsum dolor sit amet, consectetur adipiscing elit.
-    Mauris dictum quam eget orci porttitor ultrices.
-    Suspendisse purus nisl, dapibus et rhoncus in, eleifend id quam. Integer volutpat dolor urna, non auctor magna sodales non.
-    Ut in fermentum erat, lacinia malesuada urna.
-    Nullam et dolor ut purus fermentum rutrum.
-    Suspendisse non velit sollicitudin, rutrum est at, commodo sem. Pellentesque viverra tincidunt mauris ut egestas.
-    Integer vel orci sapien. Sed ut mi at lacus vehicula fermentum.
-    Integer euismod vestibulum auctor.
-    Nunc vulputate leo eget aliquam bibendum.
+    Hybrid MPI+X programming is the current standard for High Performance Computing Applications.
+    Parallelising within a process poses new challenges for efficient programs though.
+    A main problem is communicating logical parallelism in the application to the MPI Libaries like MPICH or OpenMPI.
+    The new MPI 4.0 standard assists application developers with new hints and partitioned operations to simplify the exposure of that parallelism.
+    This paper summarizes the operations in current MPI 4.0, gives examples of how to apply them and gives recommendations based on the applications structure.
 
@@
 
-% Introduction, roughly 1 page
-% todo: rework
 \section{Introduction}
 
-%vll. Wort zur Umsetzung in Bibliotheken? OpenMPI und MPICH haben es beide immerhin ...
-
-% Historisch MPI-Everywhere
-% heute MPI+OpenMP (warum)
-% Was sind die schwierigkeiten mit MPI+OpenMP
-% Paper versucht Lösungen zusammenzufassen
 Message Passing Interface is the de facto standard for scaling High Performance Computing (HPC) Workloads across multiple compute nodes.
@@
 
-This paper gives an overview of ways to expose parallelism in communication to MPI libaries to achieve maximum performance.
+This paper gives an overview of ways to expose parallelism in communication to MPI libaries to achieve optimal performance (over 2x increase for communication-intensive programs \cite{zambreLogicalParallel2021}).
 For this it will discuss mainly point-to-point communication.
@@
 
-%todo: is simple the right word? most backwards compatible?
 The simplest though maybe not most intuitive way to parallelise communication is to assign each logical parallel message its own communicator.
-Since MPI gives no implies no relative ordering on different communicators \cite{zambreLessonsLearned2022}, libraries can map messages on different communicators to different underlying message channels.
+Since MPI implies no relative ordering on different communicators \cite{zambreLessonsLearned2022}, libraries can map messages on different communicators to different underlying message channels.
 To use different communicator an MPI application developer needs to be aware of a few key problems:
 
-1) For optimal parallelisation there needs to be two communicator for each pair of threads communicating between different processes.
-%todo: insert figure, also link figure
-%todo: possibly code for that same figure so the complexity can be better judged by a reader?
-To demonstrate the need for two communicator per pair of threads imagine a problem devided into a line of Nodes.
+\begin{figure}
+    \caption{
+        For ideal communication parallelism for 9 thread processes with a 3x3 grid that communicate along an axis, the left and right communicators on each process have to be distinct to avoid dependencies between left and right connections.
+        Each unique colour represents a unique connector in the picture.
+    }
+    \Description[]{3x3 thread processes communicating along a single axis and the associated 3-point stencil oriented in parallelisation direction}
+    \includegraphics[width=0.47\textwidth]{Communicator_Line.png}
+    \label{fig:Communicator_Line}
+\end{figure}
+
+\begin{figure}
+    \caption{3-point stencil used in Figure \ref{fig:Communicator_Line}}
+    \Description[]{3-point stencil in parallelisation direction}
+    \includegraphics[width=0.23\textwidth]{3_point_stencil.png}
+    \label{fig:3pointStencil}
+\end{figure}
+
+\begin{figure}
+    \caption{C++ implementation of Figure \ref{fig:Communicator_Line}}
+    \Description[]{Code for 3x3 thread processes figure}
+    \includegraphics[width=0.47\textwidth]{Communicator_Line_CPP.png}
+    \label{fig:Communicator_Line_CPP}
+\end{figure}
+
+\textbf{1)} \textit{For optimal parallelisation there needs to be two communicator for each pair of threads communicating between different processes.}
+To demonstrate the need for two communicators per pair of threads, imagine a problem where processes communicate along a single axis as in Figure \ref{fig:Communicator_Line}.
 If only one communicator is used per pair of threads, communication between \verb|<Rank_0,Thread_5>| and \verb|<Rank_1,Thread_3>| shares its communicator with the communication between \verb|<Rank_1,Thread_5>| and \verb|<Rank_2,Thread_3>|.
-This is problematic since the messages from \verb|Rank_0| and \verb|Rank_2| to \verb|Rank_1| now run on the same communicator and with that those two exchanges are no longer independent.
+This is problematic since the messages from \verb|Rank_0| and \verb|Rank_2| to \verb|Rank_1| now run on the same communicator and the two exchanges are no longer independent.
+A sample implementation can be seen in Figure \ref{fig:Communicator_Line_CPP}.
 
-2) For large multidimensional stencils the number of communicators needed grows rapidly and an MPI program using those should consider other methods to expose parallelisation.
+\textbf{2)} \textit{For large multidimensional stencils the number of communicators needed grows rapidly and an MPI program using those should consider other methods to expose parallelisation.}
 For more complex stencils like a 27-point 3D stencil as seen in Hypre\cite{hypre2020} on a larger 64-core processor (data distributed as [4,4,4] chunks) the number of communicators needed to expose all possible communication parallelism amounts to $6 \cdot 4^2$ for sides plus $24 \cdot 4^2 - 8$ for the corner diagonal plus $24 \cdot 4^2 - 48$ for the edge diagonals, totalling $808$ communicators.
@@
 
-3) Using communicators to expose parallelism and group processes introduces performance loss or limits the portability across MPI libraries.
+\textbf{3)} \textit{Using communicators to expose parallelism and group processes introduces performance loss or limits the portability across MPI libraries.}
 Since MPI 4.0 lacks standardised hints regarding communicator usage, MPI libraries cannot differentiate between communicators used to expose parallelism and those that are used for grouping processes together.
 If a MPI library tries to parallelise across communicators, any communicator that is not used to expose parallelisation puts unneccessary strain on the library, reducing overall performance.
-%todo: find example hint in MPICH or OpenMP
-A program may use library specific hints to specify which communicators are used for what purpose, but doing so makes the program dependent on that MPI library instead of being compatible with any MPI library, at least for the optimal performance \cite{zambreLessonsLearned2022}.
+A MPI libary may provide extra hints to specify which communicators are used to expose parallelism, but using those makes the performance of the program depend on that library and makes swapping of libraries harder.
 
-4) Using communicators for irregular access patterns degrades performance.
-Imagine a taskbased application with $n$ worker threads and one main thread in each process.
-%todo: generate new graphic like "Lessons Learned on MPI+Threads" Fig. 5 but with N nodes to better demonstrate the infeasability of "Duplicating communicators per communication partner"
+\begin{figure}
+    \caption{processes with N worker and 1 main thread communicating irregularly; each worker thread needs to be able to communicate with all other main threads; Each coloured connection represents a unique communicator}
+    \Description[]{3 Nodes with 1 main thread and N worker threads, from each worker thread there is a connection to all other main threads, each connection is coloured by the thread id of the worker thread}
+    \includegraphics[width=0.47\textwidth]{Communicator_Irregular.png}
+    \label{fig:Communicator_Irregular}
+\end{figure}
+
+\textbf{4)} \textit{Using communicators for irregular access patterns degrades performance.}
+Imagine a taskbased application with $n$ worker threads and one main thread in each process as depicted in Figure \ref{fig:Communicator_Irregular}.
 All workers in one process may need to communicate with the main thread of any other process.
 In such a case the software may either generate $n$ communicator for each worker thread in a process or generate $n \cdot m$ communicator for all $n$ worker in a process and all $m$ processes.
-The latter version is infeasable for any non trivial number of processes as it strains the MPI library for the same reason large multidimensional stencils do.
+The latter version is infeasable for any non trivial number of processes as it generates way more communicators which strains the MPI library for the same reason large multidimensional stencils do.
 So the software is realistically restricted to just $n$ communicators.
@@
 This leads to communicators performing worse than other parallelisation alternatives.
-As example communicators performed 1.63x worse than tags with hints in the task-based framework Legion \cite{zambreLogicalParallel2021}.
+As an example communicators performed 1.63x worse than tags with hints in the task-based framework Legion \cite{zambreLogicalParallel2021}.
 
@@
 
-MPI operations like \verb|MPI_Send| take a \verb|<Rank,Tag,Communicator>| triple as input.
+MPI operations like \verb|MPI_SEND| take a \verb|<Rank,Tag,Communicator>| triple as input.
 The tag is included to differentiate messages between the same processes on a single communicator.
@@
 To use this new avenue to expose parallelism, programs need to use the \verb|mpi_assert_no_any_tag| hint.
+
+\begin{figure}
+    \caption{
+        To encode source and destination into a tag they can simply be appended to the left side, given there is still space.
+    }
+    \Description[]{Code to encode source and target into tag}
+    \includegraphics[width=0.47\textwidth]{Tags_CPP.png}
+    \label{fig:Tags_CPP}
+\end{figure}
+
 Now any MPI library is free to parallelise over different tags.
 To use that a program can simply encode the source thread and target thread into the tag.
-A possible implementation is given in FIG. %todo: add figure of sample code
+If the \verb|MPI_ANY_TAG| is needed for application logic, but the order of the operations getting matched is irrelevant, the program may instead use the \verb|mpi_assert_allow_overtaking| to ignore the non-overtaking behaviour completely on a communicator.
+A possible implementation is given in Figure \ref{fig:Tags_CPP}.
 Using tags comes with some caveats:
 
-As with using communicators to expose parallelism, tags are also used in other contexts than parallelisation.
-So as with communicators a program has to rely on MPI library specific extra hints to clarify which parts of the tag encodes the parallelisation information or risk a suboptimal distribution to the hardware communication channels.
-An example of such extra hints is shown in FIG for the MPI library MPICH. %todo: add MPICH figure
+As with communicators the other uses of tags may strain a library more than needed and as with communicators libraries may include extra hints to help the mapping of tags to hardware channels.
+But using those hints once again binds the program performance to a specific library and makes swapping libraries harder.
 
@@
 
-To implement partitioned point-to-point operations is slightly more difficult than simple \verb|MPI_Send| and \verb|MPI_Recv| calls.
-In contrast to those all operations have to be first initialised on sender and receiver side with \verb|MPI_PSend_Init| and \verb|MPI_PRecv_Init| respectively.
+To implement partitioned point-to-point operations is slightly more difficult than simple \verb|MPI_SEND| and \verb|MPI_RECV| calls.
+In contrast to those all operations have to be first initialised on sender and receiver side with \verb|MPI_PSEND_INIT| and \verb|MPI_PRECV_INIT| respectively.
 The initialisation may be done long before the operation is executed.
-To start the execution there must be a single call to \verb|MPI_Start|.
-Then each contributing sending thread may write its part of the message in the shared buffer and then mark its part as ready by calling \verb|MPI_Pready|.
-On the receiving side threads may poll for completion of individual partitions of the message by calling \verb|MPI_Parrived|.
+To start the execution there must be a single call to \verb|MPI_START|.
+Then each contributing sending thread may write its part of the message in the shared buffer and then mark its part as ready by calling \verb|MPI_PREADY|.
+On the receiving side threads may poll for completion of individual partitions of the message by calling \verb|MPI_PARRIVED|.
 As soon as those return a truthy flag a thread may work on that partitions data.
-There must be a single final call to \verb|MPI_Wait| or \verb|MPI_Test| to complete the operation, before it may be restarted with another call to \verb|MPI_Start|.
+There must be a single final call to \verb|MPI_WAIT| or \verb|MPI_TEST| to complete the operation, before it may be restarted with another call to \verb|MPI_START|.
 
@@
 They are very ill suited for irregular communication though, as the operations have to be initialised before any communication may happen \cite{zambreLessonsLearned2022}.
-Irregular communication is further complicated by the missing ability to provide wildcards (\verb|MPI_ANY_TAG|, \verb|MPI_ANY_SOURCE|) to \verb|MPI_PRecv_Init| \cite{mpi40}, on which modern task-based runtimes like Legion rely \cite{zambreLessonsLearned2022}.
+Irregular communication is further complicated by the missing ability to provide wildcards (\verb|MPI_ANY_TAG|, \verb|MPI_ANY_SOURCE|) to \verb|MPI_PRECV_INIT| \cite{mpi40}, on which modern task-based runtimes like Legion rely \cite{zambreLessonsLearned2022}.
 
@@
 
-MPI implements Remote Memory Access (RMA) with memory windows.
-In those windows other processes may put data with \verb|MPI_Put| or get data from with \verb|MPI_Get| or their non-blocking alternatives \cite{mpi40}.
-In contrast to point-to-point communication MPI imposes no order requirement for \verb|MPI_Put| and \verb|MPI_Get| operations \cite{mpi40}.
+Remote Memory Accesses (RMA) are one sided communication, where the target does not need to do anything after the initialisation.
+Those operations are only allowed on specific sections of their memory called windows in MPI.
+To create such a window, all processes in a communicator have to call \verb|MPI_WIN_CREATE|.
+They may specify different memory sizes for their window, including specifying \verb|size=0| to not expose any memory themselves.
+In those windows other processes may put data with \verb|MPI_PUT| or get data from with \verb|MPI_GET| or their non-blocking alternatives \cite{mpi40}.
+These RMA operations are not matched on the destination, hence being called one-sided.
+In contrast to point-to-point communication MPI imposes no order requirement for \verb|MPI_PUT| and \verb|MPI_GET| operations \cite{mpi40}.
 So for these operations MPI libraries are free to parallelise them as needed.
-MPI windows also allow atomic \verb|MPI_Accumulate| operations.
+MPI windows also allow atomic \verb|MPI_ACCUMULATE| operations.
 This operation updates a target value by adding a number to it atomically.
-MPI imposes strict ordering on \verb|MPI_Accumulate| operations per default.
+MPI imposes strict ordering on \verb|MPI_ACCUMULATE| operations per default.
 If such a ordering guarantee is not needed, Zambre and Chandramowlishwaran list NWChem as an example \cite{zambreLessonsLearned2022}, it may be relaxed on window creation by specifying another \verb|accumulate_ordering|.
@@
 \section{Conclusions}
+
+In this paper I demonstrate multiple ways to expose communication parallelism to MPI libraries in MPI 4.0 while using CPU multithreading.
+It has been demonstrated that communication can be a significant bottleneck and exposing available parallelism to MPI libraries increase performance by over 2x for some applications \cite{zambreLogicalParallel2021}.
+From the described methods I would recommend tags with hints if the application is allowed to depend on a specific MPI library and does not rely on \verb|MPI_ANY_TAG|.
+It is the easiest to include into existing applications and performs no worse than its alternatives \cite{zambreLessonsLearned2022}.
+Any of the described methods will work for most application types though.
+
 \bibliographystyle{ACM-Reference-Format}

### 2024-07-15 HPC/mpi.h

@@ -1,0 +1,1884 @@
+/* ompi/include/mpi.h.  Generated from mpi.h.in by configure.  */
+/*
+ * Copyright (c) 2004-2005 The Trustees of Indiana University and Indiana
+ *                         University Research and Technology
+ *                         Corporation.  All rights reserved.
+ * Copyright (c) 2004-2006 The University of Tennessee and The University
+ *                         of Tennessee Research Foundation.  All rights
+ *                         reserved.
+ * Copyright (c) 2004-2007 High Performance Computing Center Stuttgart,
+ *                         University of Stuttgart.  All rights reserved.
+ * Copyright (c) 2004-2005 The Regents of the University of California.
+ *                         All rights reserved.
+ * Copyright (c) 2007-2009 Cisco Systems, Inc.  All rights reserved.
+ * Copyright (c) 2008-2009 Sun Microsystems, Inc.  All rights reserved.
+ * $COPYRIGHT$
+ *
+ * Additional copyrights may follow
+ *
+ * $HEADER$
+ */
+
+#ifndef OMPI_MPI_H
+#define OMPI_MPI_H
+
+/* The comment below (and the ending partner) are for building fat
+   distributions on platforms that support it.  Please do not remove */
+
+/* @OMPI_BEGIN_CONFIGURE_SECTION@ */
+
+#ifndef OMPI_CONFIG_H
+
+/* Only include these if OMPI_CONFIG_H isn't defined (meaning if
+   ompi_config.h hasn't already been included).  Otherwise, we'll
+   duplicate all those symbols.  OMPI coding standards say that
+   ompi_config.h must be included before all other files, so this
+   should be good enough */
+
+/* Define to 1 if you have the ANSI C header files. */
+#define OMPI_STDC_HEADERS 1
+
+/* Define to 1 if you have the <sys/time.h> header file. */
+#define OMPI_HAVE_SYS_TIME_H 1
+
+/* Define to 1 if you have the <sys/synch.h> header file. */
+/* #undef OMPI_HAVE_SYS_SYNCH_H */
+
+/* Define to 1 if the system has the type `long long'. */
[…]

### 2024-07-15 HPC/src/Communicator_Line.cpp

@@ -1,0 +1,50 @@
+#include "mpi.h"
+#include "omp.h"
+
+int main(int argc, char *argv[])
+{
+    int dataLen = 100;
+    float dataLeft[dataLen], dataRight[dataLen];
+    bool done = false;
+    MPI_Init(&argc, &argv);
+    int rank;
+    MPI_Comm_rank(MPI_COMM_WORLD, &rank);
+    MPI_Comm comm_0[3], comm_1[3];
+    for (int i = 0; i < 3; i++)
+    {
+        MPI_Comm_dup(MPI_COMM_WORLD, &comm_0[i]);
+        MPI_Comm_dup(MPI_COMM_WORLD, &comm_1[i]);
+    }
+
+#pragma omp parallel num_threads(9)
+    {
+        while (!done)
+        {
+            // halo exchange
+            int tid = omp_get_thread_num();
+            int tid_x = tid % 3;
+            int tid_y = tid / 3;
+            if (tid_x == 0)
+            {
+                MPI_Comm left_comm = rank % 2 ? comm_1[tid_y] : comm_0[tid_y];
+                MPI_Sendrecv_replace(&dataLeft, dataLen, MPI_FLOAT, rank - 1, 0, rank - 1, 0, left_comm, MPI_STATUS_IGNORE);
+            }
+            else if (tid_x == 2)
+            {
+                MPI_Comm right_comm = rank % 2 ? comm_0[tid_y] : comm_1[tid_y];
+                MPI_Sendrecv_replace(&dataRight, dataLen, MPI_FLOAT, rank + 1, 0, rank + 1, 0, right_comm, MPI_STATUS_IGNORE);
+            }
+            // update boundaries  with new data
+            // calculate next step
+        }
+    }
+
+    for (int i = 0; i < 3; i++)
+    {
+        MPI_Comm_free(&comm_0[i]);
+        MPI_Comm_free(&comm_1[i]);
+    }
+
+    MPI_Finalize();
+    return 0;
+}

### 2024-07-15 HPC/paper.tex
Added Sample Codes to RMA and Partitioned Operations

@@
 
-% todo: revisit?
 \begin{abstract}
@@
     Parallelising within a process poses new challenges for efficient programs though.
-    A main problem is communicating logical parallelism in the application to the MPI Libaries like MPICH or OpenMPI.
+    A main problem is communicating logical parallelism in the application to the MPI libraries like MPICH or OpenMPI.
     The new MPI 4.0 standard assists application developers with new hints and partitioned operations to simplify the exposure of that parallelism.
@@
 This paper will speak of MPI as the standard and refer to MPI implementations like MPICH and OpenMPI as "MPI libraries" going forward.
-These MPI libaries implement the MPI standard while trying to achieve maximum performance while adhearing to it.
+These MPI libraries implement the MPI standard while trying to achieve maximum performance while adhearing to it.
 To do so some libraries include extra library specific settings as hints, which the library user can set to increase possible performance.
-Sometimes these libraries will also include entirely new ways to parallelise communication like MPIX\_Stream \cite{Zhou2022}, but this paper will focus mainly on the MPI standard itself.
+Sometimes these libraries will also include entirely new ways to parallelize communication like MPIX\_Stream \cite{Zhou2022}, but this paper will focus mainly on the MPI standard itself.
 
 Traditionally applications relied on the memory-hungry MPI-Everywhere model \cite{zambreLessonsLearned2022}.
-In the last decades however a rapid increase in core counts per CPU relative to other on-node resources neccessitated a switch to a model of MPI+X,
-where MPI is used to parallise between different nodes and X is used to parallelise on each node.
+In the last decades however a rapid increase in core counts per CPU relative to other on-node resources necessitated a switch to a model of MPI+X,
+where MPI is used to parallelize between different nodes and X is used to parallelize on each node.
 Some popular choices for this second X part are CPU multithreading libraries like OpenMP or POSIX Threads, GPU Acceleration with CUDA or using MPI with Remote Memory Access calls.
@@
 Its ease of use also makes it popular for developing HPC software.
-MPI+OpenMP programs tend to perform worse than their MPI-Everywhere counterparts \cite{zambreLessonsLearned2022}\cite{zambreLogicalParallel2021}.
+MPI+OpenMP programs tend to perform worse than their MPI-Everywhere counterparts \cite{zambreLessonsLearned2022,zambreLogicalParallel2021}.
 The main problem being extra unneeded thread synchronization overheads \cite{zambreLessonsLearned2022} and less exposure of parallelism for MPI libraries to use.
 
-This paper gives an overview of ways to expose parallelism in communication to MPI libaries to achieve optimal performance (over 2x increase for communication-intensive programs \cite{zambreLogicalParallel2021}).
+This paper gives an overview of ways to expose parallelism in communication to MPI libraries to achieve optimal performance (over 2x increase for communication-intensive programs \cite{zambreLogicalParallel2021}).
 For this it will discuss mainly point-to-point communication.
@@
 The historically preferred method of MPI-Everywhere made logical parallelism easy to express in software.
-It does however use up more memory and other on chip ressources relative to MPI+OpenMP.
+It does however use up more memory and other on chip resources relative to MPI+OpenMP.
 Classical MPI-Everywhere programs spawn a single process per core available.
@@
 Each MPI process also needs to be a full program, which means each core has to have all instructions duplicated.
-OpenMP however only needs to generate code for the actually parallised parts (typically loops) of the program, which reduces the size for all additional cores significantly.
+OpenMP however only needs to generate code for the parallized parts (typically loops) of the program, which reduces the size for all additional cores significantly.
 
@@
 
-% Body, roughly 2.5 pages
 \section{Parallel Point-to-Point Communication in MPI 4.0}
@@
 MPI 4.0 also added new hints, a way for a program to guarantee MPI that it will not use certain features, that can make tags viable for performant parallel communication.
-Furthermore MPI 4.0 introduced MPI Sessions, which may be used for parallelisation purposes, but are not considered in the scope of this paper.
+Furthermore MPI 4.0 introduced MPI Sessions, which may be used for parallelization purposes, but are not considered in the scope of this paper.
 In the following subsections this paper will discuss the mentioned methods in more detail.
@@
 
-The simplest though maybe not most intuitive way to parallelise communication is to assign each logical parallel message its own communicator.
+The simplest though maybe not most intuitive way to parallelize communication is to assign each logical parallel message its own communicator.
 Since MPI implies no relative ordering on different communicators \cite{zambreLessonsLearned2022}, libraries can map messages on different communicators to different underlying message channels.
@@
     \caption{3-point stencil used in Figure \ref{fig:Communicator_Line}}
-    \Description[]{3-point stencil in parallelisation direction}
+    \Description[]{3-point stencil in parallelization direction}
     \includegraphics[width=0.23\textwidth]{3_point_stencil.png}
@@
 
-\textbf{1)} \textit{For optimal parallelisation there needs to be two communicator for each pair of threads communicating between different processes.}
+\textbf{1)} \textit{For optimal parallelization there needs to be two communicator for each pair of threads communicating between different processes.}
 To demonstrate the need for two communicators per pair of threads, imagine a problem where processes communicate along a single axis as in Figure \ref{fig:Communicator_Line}.
@@
 
-\textbf{2)} \textit{For large multidimensional stencils the number of communicators needed grows rapidly and an MPI program using those should consider other methods to expose parallelisation.}
+\textbf{2)} \textit{For large multidimensional stencils the number of communicators needed grows rapidly and an MPI program using those should consider other methods to expose parallelization.}
 For more complex stencils like a 27-point 3D stencil as seen in Hypre\cite{hypre2020} on a larger 64-core processor (data distributed as [4,4,4] chunks) the number of communicators needed to expose all possible communication parallelism amounts to $6 \cdot 4^2$ for sides plus $24 \cdot 4^2 - 8$ for the corner diagonal plus $24 \cdot 4^2 - 48$ for the edge diagonals, totalling $808$ communicators.
@@
 Since MPI 4.0 lacks standardised hints regarding communicator usage, MPI libraries cannot differentiate between communicators used to expose parallelism and those that are used for grouping processes together.
-If a MPI library tries to parallelise across communicators, any communicator that is not used to expose parallelisation puts unneccessary strain on the library, reducing overall performance.
+If a MPI library tries to parallelise across communicators, any communicator that is not used to expose parallelisation puts unnecessary strain on the library, reducing overall performance.
 A MPI libary may provide extra hints to specify which communicators are used to expose parallelism, but using those makes the performance of the program depend on that library and makes swapping of libraries harder.
@@
 \textbf{4)} \textit{Using communicators for irregular access patterns degrades performance.}
-Imagine a taskbased application with $n$ worker threads and one main thread in each process as depicted in Figure \ref{fig:Communicator_Irregular}.
+Imagine a task-based application with $n$ worker threads and one main thread in each process as depicted in Figure \ref{fig:Communicator_Irregular}.
 All workers in one process may need to communicate with the main thread of any other process.
 In such a case the software may either generate $n$ communicator for each worker thread in a process or generate $n \cdot m$ communicator for all $n$ worker in a process and all $m$ processes.
-The latter version is infeasable for any non trivial number of processes as it generates way more communicators which strains the MPI library for the same reason large multidimensional stencils do.
+The latter version is infeasible for any non trivial number of processes as it generates way more communicators which strains the MPI library for the same reason large multidimensional stencils do.
 So the software is realistically restricted to just $n$ communicators.
@@
 Normally the MPI standard requires a non-overtaking order for all messages with the same \verb|<Rank,Communicator>| \cite{mpi40}.
-So the default settings do not enable a program to use the additionally sent tags for parallelisation.
+So the default settings do not enable a program to use the additionally sent tags for parallelization.
 If the program however guarantees MPI, that it will not use the \verb|MPI_ANY_TAG| (a parameter matching all tags), MPI libraries can send different messages between the same ranks but differing tags in arbitrary order.
@@
     \caption{
-        To encode source and destination into a tag they can simply be appended to the left side, given there is still space.
+        C++ Code to encode source and destination into a tag.
+        The bitpattern of source and destination are simply appended at the left side of the tag.
     }
@@
 
-Now any MPI library is free to parallelise over different tags.
+Now any MPI library is free to parallelize over different tags.
 To use that a program can simply encode the source thread and target thread into the tag.
@@
 For a 128 core processors this would mean, that a program reserves $2\log_2{128} = 14$ bits of their tags.
-For most programs this will be irrelevant, but if the program already encodes much information in their tags, exposing parallelisation information might not be feasable.
+For most programs this will be irrelevant, but if the program already encodes much information in their tags, exposing parallelisation information might not be feasible.
 
@@
 
-%todo: add sample code
+\begin{figure}
+    \caption{
+        A C++ example of how to use partitioned operations with MPI.
+        In the example 2 MPI processes do some calculation together, where the output of the first process gets consumed by the second process.
+        There exists some synchronization to ensure only a single thread is able to call MPI\_Start and MPI\_Wait.
+    }
+    \Description[]{C++ Example Code for partitioned operations.}
+    \includegraphics[width=0.47\textwidth]{PartitionOperations_CPP.png}
+    \label{fig:PartitionedOps_CPP}
+\end{figure}
 
-Partitioned Operations are point-to-point operations where the message is partioned into multiple parts.
+Partitioned Operations are point-to-point operations where the message is partitioned into multiple parts.
 
@@
 There must be a single final call to \verb|MPI_WAIT| or \verb|MPI_TEST| to complete the operation, before it may be restarted with another call to \verb|MPI_START|.
+An example of this is visible in Figure \ref{fig:PartitionedOps_CPP}
 
@@
 
+\begin{figure}
+    \caption{
+        A C++ example of how to create a MPI\_Window, that does not enforce ordering for MPI\_Accumulate operations.
+    }
+    \Description[]{C++ Example Code to create a MPI_Window without ordering on accumulate}
+    \includegraphics[width=0.47\textwidth]{RMANoOrdering.png}
+    \label{fig:RMANoOrdering_CPP}
+\end{figure}
+
 Remote Memory Accesses (RMA) are one sided communication, where the target does not need to do anything after the initialisation.
@@
 In contrast to point-to-point communication MPI imposes no order requirement for \verb|MPI_PUT| and \verb|MPI_GET| operations \cite{mpi40}.
-So for these operations MPI libraries are free to parallelise them as needed.
+So for these operations MPI libraries are free to parallelize them as needed.
 MPI windows also allow atomic \verb|MPI_ACCUMULATE| operations.
@@
 The MPI standard supports specifying no ordering or only a subset of read-after-read, read-after-write, write-after-read, write-after-write in such a case \cite{mpi40}.
-%todo: add example with no ordering
+An example of no ordering can be seen in Figure \ref{fig:RMANoOrdering_CPP}.
 
-% End, roughly 0.5 pages
 \section{Conclusions}

### 2024-07-15 HPC/src/Communicator_Line.cpp

@@ -1,4 +1,4 @@
-#include "mpi.h"
+#include "../mpi.h"
 #include "omp.h"
 
 int main(int argc, char *argv[])

### 2024-07-15 HPC/src/partitioned_Operations.cpp

@@ -1,0 +1,51 @@
+// example from MPI 4.0 standard https://www.mpi-forum.org/docs/mpi-4.0/mpi40-report.pdf
+#include "../mpi.h"
+int main()
+{
+    // initialize MPI and get own rank as rank
+    float *buffer = new float[numThreads * arrLengthPerThread];
+    bool done = false;
+    MPI_Request request;
+    if (rank == 0)
+    {
+        MPI_Psend_init(buffer, numThreads, arrLengthPerThread, MPI_FLOAT, 0, 0, MPI_COMM_WORLD, info, &request);
+#pragma omp parallel shared(request) num_threads(numThreads)
+        while (!done)
+        {
+#pragma omp single
+            MPI_Start(&request);
+#pragma omp for
+            for (int i = 0; i < numThreads; i++)
+            {
+                /* compute and fill partition #i, then mark ready: */
+                MPI_Pready(i, request);
+            }
+#pragma omp single
+            MPI_Wait(&request, &flag, MPI_STATUS_IGNORE);
+        }
+    }
+    else if (rank == 1)
+    {
+        MPI_Precv_init(buffer, numThreads, arrLengthPerThread, MPI_FLOAT, 0, 0, info, MPI_COMM_WORLD, &request);
+#pragma omp parallel shared(request) num_threads(numThreads)
+        while (!done)
+        {
+            int arrived = 0;
+#pragma omp single
+            MPI_Start(&request);
+#pragma omp for
+            for (int i = 0; i < numThreads; i++)
+            {
+                while (arrived == 0)
+                {
+                    MPI_Parrived(i, request, &arrived);
+                }
+                // process partition #i
+            }
+            MPI_Wait(&request, &flag, MPI_STATUS_IGNORE);
+        }
+    }
+    MPI_Request_free(&request);
+    MPI_Finalize();
+    return 0;
+}

### 2024-07-15 HPC/src/unordered_RMA.cpp

@@ -1,0 +1,21 @@
+#include "../mpi.h"
+#include <string>
+
+int main(int argc, char *argv[])
+{
+    MPI_Init(&argc, &argv);
+    int rank;
+    MPI_Comm_rank(MPI_COMM_WORLD, &rank);
+    float buffer[100];
+
+    MPI_Info info;
+    MPI_Info_create(&info);
+    std::string key = "accumulate_ordering";
+    std::string value = "none";
+    MPI_Info_set(info, key.data(), value.data());
+
+    MPI_Win win;
+    MPI_Win_create(buffer, 100 * sizeof(float), sizeof(float), info, MPI_COMM_WORLD, &win);
+
+    return 0;
+}

### 2024-07-15 HPC/paper.tex
Final Version of paper

@@
     Hybrid MPI+X programming is the current standard for High Performance Computing Applications.
-    Parallelising within a process poses new challenges for efficient programs though.
+    This paper focuses on how to write performant MPI+OpenMP programs.
+    Parallelising with OpenMP and MPI together poses new challenges for efficient programs though.
     A main problem is communicating logical parallelism in the application to the MPI libraries like MPICH or OpenMPI.
     The new MPI 4.0 standard assists application developers with new hints and partitioned operations to simplify the exposure of that parallelism.
-    This paper summarizes the operations in current MPI 4.0, gives examples of how to apply them and gives recommendations based on the applications structure.
+    This paper summarizes the operations that support fine grained parallelisation in current MPI 4.0, gives examples of how to apply them and gives recommendations based on the applications structure.
 
@@
 
-Message Passing Interface is the de facto standard for scaling High Performance Computing (HPC) Workloads across multiple compute nodes.
+Message Passing Interface (MPI) is the de facto standard for scaling High Performance Computing (HPC) Workloads across multiple compute nodes.
 This paper will speak of MPI as the standard and refer to MPI implementations like MPICH and OpenMPI as "MPI libraries" going forward.
-These MPI libraries implement the MPI standard while trying to achieve maximum performance while adhearing to it.
+These MPI libraries implement the MPI standard while trying to achieve maximum performance.
 To do so some libraries include extra library specific settings as hints, which the library user can set to increase possible performance.
-Sometimes these libraries will also include entirely new ways to parallelize communication like MPIX\_Stream \cite{Zhou2022}, but this paper will focus mainly on the MPI standard itself.
+Sometimes these libraries will also include entirely new ways to parallelize communication like MPIX\_Stream in MPICH \cite{Zhou2022}, but this paper will focus on methods in the MPI standard itself.
 
@@
 where MPI is used to parallelize between different nodes and X is used to parallelize on each node.
-Some popular choices for this second X part are CPU multithreading libraries like OpenMP or POSIX Threads, GPU Acceleration with CUDA or using MPI with Remote Memory Access calls.
+Some popular choices for this second X part are CPU multithreading libraries like OpenMP or POSIX Threads, GPU Acceleration with CUDA or using MPI with Remote Memory Access calls for the communication within a node.
 This paper will focus on MPI+OpenMP, though many aspects of this paper also apply for any other CPU multithreading library.
@@
 Its ease of use also makes it popular for developing HPC software.
-MPI+OpenMP programs tend to perform worse than their MPI-Everywhere counterparts \cite{zambreLessonsLearned2022,zambreLogicalParallel2021}.
+Currently MPI+OpenMP programs tend to perform worse than their MPI-Everywhere counterparts \cite{zambreLessonsLearned2022,zambreLogicalParallel2021}.
 The main problem being extra unneeded thread synchronization overheads \cite{zambreLessonsLearned2022} and less exposure of parallelism for MPI libraries to use.
@@
 It does however use up more memory and other on chip resources relative to MPI+OpenMP.
-Classical MPI-Everywhere programs spawn a single process per core available.
-MPI+OpenMP programs normally spawn a single process per node or NUMA domain \cite{zambreLessonsLearned2022}.
+Classical MPI-Everywhere programs spawn a single process per core available while MPI+OpenMP programs normally only spawn a single process per node or NUMA domain \cite{zambreLessonsLearned2022}.
 
@@
 Assuming an $n$ core processor $n$ copies of the list are lying in each nodes memory in the MPI-Everywhere model.
-An equivalent MPI+OpenMP program shares its list between its cores and also spawns $n$ times less processes, so \verb|MPI_COMM_WORLD| is $n$ times smaller and $n$ times less often on each processor.
+An equivalent MPI+OpenMP program shares its list between its cores and also spawns $n$ times less processes, so \verb|MPI_COMM_WORLD| is $n$ times smaller and lies $n$ times less often in the memory of each processor.
 For current generation processors exceeding 100 cores like Intel® Xeon® 6 \cite{intelXeon6} or 4th Generation AMD EPYC™ \cite{amd4thGenEpyc} this equates to a reduction of \verb|MPI_COMM_WORLD|'s memory consumption of over $n^2 > 10.000$ times.
@@
 Each MPI process also needs to be a full program, which means each core has to have all instructions duplicated.
-OpenMP however only needs to generate code for the parallized parts (typically loops) of the program, which reduces the size for all additional cores significantly.
+OpenMP however only needs to generate code for the parallized parts (typically loops) of the program, which reduces the memory consumption for all additional cores significantly.
 
@@
 This is important to prevent deadlocks (two processes each wait on the other) but also slows down the communication overall.
-We will explore how to circumvent such slowdowns in the following section.
+The rest of the paper dedicated to explain how such slowdowns may be prevented.
+The paper will assume that \verb|MPI_INIT| has been called with \verb|MPI_THREAD_MULTIPLE| to enable multithreaded access to MPI \cite{mpi40}.
 
@@
 MPI 4.0 introduced Partitioned Operations to aid efficient parallel communication.
-In Partitioned Operations multiple threads each contribute part of a larger message.
+In Partitioned Operations multiple threads each contribute a part of a larger message.
 MPI 4.0 also added new hints, a way for a program to guarantee MPI that it will not use certain features, that can make tags viable for performant parallel communication.
-Furthermore MPI 4.0 introduced MPI Sessions, which may be used for parallelization purposes, but are not considered in the scope of this paper.
+Furthermore MPI 4.0 introduced MPI Sessions, which may be used for parallelization purposes, but are not considered in the scope of this paper \cite{MPISessions}.
 In the following subsections this paper will discuss the mentioned methods in more detail.
@@
 \begin{figure}
-    \caption{3-point stencil used in Figure \ref{fig:Communicator_Line}}
-    \Description[]{3-point stencil in parallelization direction}
-    \includegraphics[width=0.23\textwidth]{3_point_stencil.png}
-    \label{fig:3pointStencil}
-\end{figure}
-
-\begin{figure}
     \caption{C++ implementation of Figure \ref{fig:Communicator_Line}}
@@
 \textbf{1)} \textit{For optimal parallelization there needs to be two communicator for each pair of threads communicating between different processes.}
-To demonstrate the need for two communicators per pair of threads, imagine a problem where processes communicate along a single axis as in Figure \ref{fig:Communicator_Line}.
+To demonstrate the need for two communicators per pair of threads, imagine a problem where processes only need to communicate along a single axis as in Figure \ref{fig:Communicator_Line}.
 If only one communicator is used per pair of threads, communication between \verb|<Rank_0,Thread_5>| and \verb|<Rank_1,Thread_3>| shares its communicator with the communication between \verb|<Rank_1,Thread_5>| and \verb|<Rank_2,Thread_3>|.
@@
 \textbf{2)} \textit{For large multidimensional stencils the number of communicators needed grows rapidly and an MPI program using those should consider other methods to expose parallelization.}
-For more complex stencils like a 27-point 3D stencil as seen in Hypre\cite{hypre2020} on a larger 64-core processor (data distributed as [4,4,4] chunks) the number of communicators needed to expose all possible communication parallelism amounts to $6 \cdot 4^2$ for sides plus $24 \cdot 4^2 - 8$ for the corner diagonal plus $24 \cdot 4^2 - 48$ for the edge diagonals, totalling $808$ communicators.
+For more complex stencils like a 27-point 3D stencil as seen in Hypre\cite{hypre2020} on a 64-core processor (data distributed as [4,4,4] chunks) the number of communicators needed to expose all possible communication parallelism amounts to $6 \cdot 4^2$ for the sides plus $24 \cdot 4^2 - 8$ for the corner diagonal plus $24 \cdot 4^2 - 48$ for the edge diagonals, totalling $808$ communicators.
 The workload would only need a single parallel communication channel for each thread communicating between nodes ideally though, which is only $4^3 - (4-2)^3 = 56$.
 The equations were derived from those used by Zambre and Chandramowlishwaran \cite{zambreLessonsLearned2022}.
-Generating so many communicators is not only complex, it also strains MPI libraries that have to map the many possibly parallel messages to a far smaller number of parallel paths in hardware (as an example Omni-Path only has 160 hardware contexts \cite{intelOmniPath}).
+Generating so many communicators is not only complex, it also strains MPI libraries that have to map the many possibly parallel messages to a far smaller number of parallel paths in hardware, (as an example Omni-Path only has 160 hardware contexts \cite{intelOmniPath}).
 Since the MPI library lacks the knowledge of the intended communication pattern its mapping to the relevant hardware contexts will be suboptimal.
@@
 This then forces the main thread to communicate over the same communicators as its workers.
-And as the main thread needs to iterate over all possible communicators it will contend with its workers to access the communicators and slow down the application that way.
+And as the main thread needs to iterate over all possible communicators it will contend with its workers to access the communicators and slow down the application.
 This leads to communicators performing worse than other parallelisation alternatives.
@@
 
-To use this new avenue to expose parallelism, programs need to use the \verb|mpi_assert_no_any_tag| hint.
-
 \begin{figure}
@@
 
+To use this new avenue to expose parallelism, programs need to use the \verb|mpi_assert_no_any_tag| hint.
 Now any MPI library is free to parallelize over different tags.
 To use that a program can simply encode the source thread and target thread into the tag.
-If the \verb|MPI_ANY_TAG| is needed for application logic, but the order of the operations getting matched is irrelevant, the program may instead use the \verb|mpi_assert_allow_overtaking| to ignore the non-overtaking behaviour completely on a communicator.
+If the \verb|MPI_ANY_TAG| is needed for application logic, but the order of the operations getting matched is irrelevant, the program may instead use the \verb|mpi_assert_allow_overtaking| to ignore the non-overtaking behaviour completely on a communicator \cite{mpi40}.
 A possible implementation is given in Figure \ref{fig:Tags_CPP}.
@@
 
-As with communicators the other uses of tags may strain a library more than needed and as with communicators libraries may include extra hints to help the mapping of tags to hardware channels.
-But using those hints once again binds the program performance to a specific library and makes swapping libraries harder.
+As with communicators the other uses of tags may strain a library more than needed.
+Libraries may include extra hints to help the mapping of tags to hardware channels, but using those hints binds the program performance to a specific library and makes swapping libraries harder.
 
@@
 For a 128 core processors this would mean, that a program reserves $2\log_2{128} = 14$ bits of their tags.
-For most programs this will be irrelevant, but if the program already encodes much information in their tags, exposing parallelisation information might not be feasible.
+For most programs this will be irrelevant, but if the program already encodes much information in their tags, exposing parallelisation information additionally in tags might not be feasible.
 
@@
 There must be a single final call to \verb|MPI_WAIT| or \verb|MPI_TEST| to complete the operation, before it may be restarted with another call to \verb|MPI_START|.
-An example of this is visible in Figure \ref{fig:PartitionedOps_CPP}
+An example of this can be seen in Figure \ref{fig:PartitionedOps_CPP}.
 
-This more complicated setup allows an MPI library to transmit the partitions in parallel instead of waiting for all threads to finish working.
+This more complicated setup allows a MPI library to transmit the partitions in parallel instead of waiting for all threads to finish working.
 Partitioned operations are ideally suited to very regular communication, as found in stencil applications.
@@
 
-Remote Memory Accesses (RMA) are one sided communication, where the target does not need to do anything after the initialisation.
+Remote Memory Accesses (RMA) are one sided communications, where the target does not need to do anything after the initialisation.
 Those operations are only allowed on specific sections of their memory called windows in MPI.
@@
 MPI imposes strict ordering on \verb|MPI_ACCUMULATE| operations per default.
-If such a ordering guarantee is not needed, Zambre and Chandramowlishwaran list NWChem as an example \cite{zambreLessonsLearned2022}, it may be relaxed on window creation by specifying another \verb|accumulate_ordering|.
+If such a ordering guarantee is not needed, Zambre and Chandramowlishwaran list NWChem as such an example \cite{zambreLessonsLearned2022}, it may be relaxed on window creation by specifying another \verb|accumulate_ordering|.
 The MPI standard supports specifying no ordering or only a subset of read-after-read, read-after-write, write-after-read, write-after-write in such a case \cite{mpi40}.
@@
 It has been demonstrated that communication can be a significant bottleneck and exposing available parallelism to MPI libraries increase performance by over 2x for some applications \cite{zambreLogicalParallel2021}.
-From the described methods I would recommend tags with hints if the application is allowed to depend on a specific MPI library and does not rely on \verb|MPI_ANY_TAG|.
+From the described methods I would recommend tags with hints if the application does not rely on the \verb|MPI_ANY_TAG|.
 It is the easiest to include into existing applications and performs no worse than its alternatives \cite{zambreLessonsLearned2022}.

### 2024-07-16 other/noten.py
Adjusted marks to order in friedolin

@@ -17,10 +17,6 @@ noten = [
     (1.0,6), # Diskrete Strukturen II 
     (1.0,6), # Einführung in die Wahrscheinlichkeitstheorie 
     (1.3,6), # Numerische Mathematik
-    # C
-    (1.3,6), # IT-Sicherheit 
-    (2.7,6), # Algorithmisches Beweisen
-    (1.3,3), # Projektmanagement (ASQ) 
     # A
     (1.0,9), # Grundlagen informatischer Problemlösung 
     (1.0,5), # Objektorientierte Programmierung 
@@ -35,9 +31,13 @@ noten = [
     (1.3,6), # Grundlagen der Informations- und Softwaresysteme 
     (1.3,4), # Kryptologie LAB
     (1.0,6), # Kryptologie
+    (2.3,6), # Einführung in die Künstliche Intelligenz
     (1.3,6), # Projekt - Paralleles Rechnen 
     (1.0,6), # Parallel Computing I  
-    (2.3,6), # Einführung in die Künstliche Intelligenz
+    # C
+    (1.3,6), # IT-Sicherheit 
+    (2.7,6), # Algorithmisches Beweisen
+    (1.3,3), # Projektmanagement (ASQ) 
     ]
 
 (avg, ects) = calcSchnitt(noten)

### 2024-07-17 other/noten.py
Added Physik Note

@@ -38,6 +38,7 @@ noten = [
     (1.3,6), # IT-Sicherheit 
     (2.7,6), # Algorithmisches Beweisen
     (1.3,3), # Projektmanagement (ASQ) 
+    (2.3,6), # Mathematische Methoden der Physik I
     ]
 
 (avg, ects) = calcSchnitt(noten)
